### 2014-02-05 doc/sfincsPaper/refereeResponse/sfincsPaper_refereeResponse.tex
Started referee response

\documentclass[11pt]{article}
\usepackage[dvips,usenames,dvipsnames]{color}
\usepackage[dvips]{graphicx}
\usepackage{delarray}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\setlength {\parindent} { 10mm} 
\setlength{\textheight}{240mm} 
\setlength{\textwidth}{160mm} 
\setlength{\oddsidemargin}{0mm}
\setlength{\topmargin}{-10mm}

% newcommands
\newcommand{\vect}[1]{\mbox{\boldmath $#1$}}
\newenvironment{referee}{\begin{quote}\it\color{Blue}}{\end{quote}}

\begin{document}  
\section*{Response to the referee}

Re: ``Comparison of particle trajectories and collision operators for
collisional transport in nonaxisymmetric plasmas'', by
M Landreman et al., manuscript POP42544.\newline

Thank you for your careful reading of the manuscript and constructive suggestions.
Below we address each of your points in detail.

\begin{referee}
If it is not difficult numerically to retain the poloidal and toroidal magnetic drifts in the
equations of motion, as it is written in the middle of page 13, why do authors keep these term
and remove only the component from eqs.(12) and (14)? I understand that it is to make
the ``DKES trajectories'', but it seems that approximating in (12) and (14) still
holds the conservation of $\mu$. Is it just because ``The omitted terms may be important in other
situations, but here our primary interest is the treatment of the $d\Phi/d\psi$ terms'' that the magnetic
drift term is completely dropped from (12) and (14), or is there any other unexplained problems
in solving the transport equation with the poloidal and toroidal magnetic drift terms kept?
\end{referee}

Response goes here.

\begin{referee}
Here the conservation properties of mass and energy in three trajectory models are discussed,
but how about the parallel momentum balance? Relating to this question, in page 20 - 21
authors have discussed about the relation between the momentum balance and friction force
from different collision operator models. I agree with the authors that the momentum
conservation property in collision operator is important to evaluate correct transport coefficients.
However, by seeing the figures 3(c)(d), parallel momentum balance seems to be broken in the
partial trajectory model, even if the full linearized Fokker-Planck operator is used.
The momentum balance equation (36) is derived by taking the $\vect{b}\cdot \int d^3v\, m\vect{v}$ moment of the drift
kinetic equation (15). Then, for each three trajectory model, taking the $\vect{b}\cdot \int d^3v\, m\vect{v}$ moment of
eq. (15) with combining one of eq. (17),(18), or (19), does it yield the proper parallel momentum
balance equation as eq.(36)? Or more simply, do these three trajectory models evaluate the
parallel viscosity $\left<\vect{b}\cdot\nabla\cdot\vect{\pi}\right>$ correctly?
\end{referee}

Response goes here.

\begin{referee}
I do not understand why the source term (25) does not work well with the pitch-angle scattering
operator or P.-A. scattering + momentum conservation operator. It is not the difference in the
collision operator but difference in the trajectory model that is related to the conservation
property of energy, (22). If the original motivation of introducing the term is to maintain the
mass and energy conservation property for finite-Er, then why you need to change the form of
source/sink term if you choose P.-A. scattering operator, which actually conserves the energy,
while you do not need according to the trajectory model? Does it mean diffusion process in
v-coordinate is essential for the source/sink term to work properly?
Please explain the reason of changing the source term according to the collision operator more
detail.
\end{referee}

Response goes here.

\begin{referee}
Concerning the source/sink term, I think that the usage of such an artificial term is relevant if
the effect of the term is small. Can you show the effect of the source/sink term is small or not in
some cases shown in Section VI, by comparing $\left< \int d^3v\, S(\psi,v) \right> \Delta t$ and $n(\psi)$, for example, where
$\Delta t$ is some characteristic time scale such as collision time or transport time scale.
\end{referee}

Response goes here.

\begin{referee}
Explanation for Fig 1 and 2: Though authors write ``As $E_r \to 0$, all the matrix elements converge
smoothly to their $E_r = 0$ limits'', since the horizontal axis of these figures are in log scale, we
cannot judge if their claim is true or not. Please show the value of $L_{jk}(E_r=0)$ in these figures,
too.
\end{referee}

Response goes here.

\begin{referee}
It is estimated that $E_* \sim 0.1 - 0.01$ in most of the plasma. However, if the core $T_e$ is higher than
$T_i$ and ambipolar $E_r$ is positive (electron root) in ECH-heated plasma, $E_*$ for ions easily becomes
nearly, or even larger than, unity near the magnetic axis, since $\rho_\theta \propto 1/B_p$ is large and usually
$|E_r|$(ion root) $\ll E_r$(electron root). See Yokoyama et al., Nuclear Fusion (2007), 1213 for example.
I think therefore the statement ``In most of the plasma, $E_*$ is however expected to be smaller
than a few percent.'' is too strong and may be misleading.
\end{referee}

Response goes here.

\begin{referee}
Appendix: The authors claims that the parallel flow and heat flux solved by SFINCS code obeys the
isomorphism, but is radial electric field ($d\Phi/dr$) contained in ``other input quantities''? Please
clarity that. Also, if the isomorphism is satisfied for finite $d\Phi/dr$ and finite source term, I
suppose that the usage of source term in the full trajectory model is justified at least in
quasisymmetric systems. Can author comment on this point in Appendix?
\end{referee}

Response goes here.

\vspace{0.3in}

In addition to the aforementioned changes, the following modifications have been made to the manuscript:

\begin{itemize}
\item Albert Moll\'{e}n has recently demonstrated precise agreement between SFINCS and an anlytical calculation
of Simakov and Helander for the limit of high collisionality [43]. His results are displayed in a new figure 6, with details given in Appendix B,
 and he has been included as a co-author.
\item Output from the DKES code is now plotted in figure 5, to demonstrate the good agreement with SFINCS.
\end{itemize}

Sincerely yours,\newline

M. Landreman

H. M. Smith

A. Moll\'{e}n

P. Helander

\end{document}