### 2015-05-21 doc/manual/overview.tex

\chapter{Overview}

SFINCS is a nifty code \cite{sfincsPaper}.

\section{\sfincs~vs. \sfincsScan}
The fortran part of \sfincs~ solves the kinetic equation for each species
at a single flux surface, a single value of $E_r$, and a single set of resolution parameters.
However, often the goal is to determine the ambipolar $E_r$ at one or more surfaces.
For this task, the \sfincsScan~family of python scripts

\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=1.0\textwidth]{poloidal_cross-sections.pdf}
\mycaption{Trajectories of trapped particles.  (a) In tokamaks and omnigenous stellarators,
all collisionless trajectories are confined.  (b) In contrast, trapped particles in other stellarators
may have a radial drift that does not average to zero.
\label{fig:poloidal_cross_sections}}
\end{center}
\end{figure}