### 2015-05-21 doc/manual/resolution.tex

\chapter{Numerical resolution parameters}

\section{Convergence testing}

\section{Typical resolution requirements}