### 2015-05-22 doc/manual/overview.tex

\chapter{Overview}

The \sfincs~code is a free, open-source tool for solving neoclassical-type kinetic problems in nonaxisymmetric or axisymmetric plasmas
with nested toroidal flux surfaces.
The code solves a drift-kinetic equation for an arbitrary number of species.  Optionally, a quasi-neutrality
equation can also be solved to obtain the self-consistent variation of the electrostatic potential on a flux surface.

This manual describes ``version 3'' of \sfincs.  There are two older versions of the
code called singleSpecies and multiSpecies.

This document discusses use and operation of the code.  For more details about the specific equations implemented,
see the version 3 technical documentation available in the \path{sfincs/docs} directory.
Ref \cite{sfincsPaper} gives many details and some early physics results.

\section{GMRES/KSP and preconditioning}

At its heart, \sfincs~solves one or more large sparse linear systems $Ax=b$.  Here $b$ is a known right-hand side vector,
$A$ is a large (often millions $\times$ millions) known sparse matrix, and $x$ is the desired and unknown solution vector.
The direct way to solve such systems is to $LU$-factorize
the matrix $A$ into lower- and upper-triangular factors.
Once the $L$ and $U$ factors are found, the solution of the linear system
for any right-hand side vector can be rapidly obtained.  However, even if the original matrix is sparse, the $L$ and $U$ factors
are generally not sparse, and so a very large amount of memory can be required for a direct $LU$-factorization.

An alternative way to solve
such large linear systems is with a so-called ``Krylov-space'' iterative method, which can dramatically reduce the memory required
compared to a direct solution.  For the non-symmetric matrices that arise in \sfincs, the preferred Krylov-space
algorithm is called GMRES (Generalized Minimal RESidual.)  In \sfincs, the \PETSc~library is used to solve the large systems
of equations. \PETSc~calls its family of linear solvers KSP, so in the output of \sfincs~you will see a ``KSP residual'' reported
as GMRES iterates towards the solution.

An important element of Krylov methods is preconditioning.  The art of preconditioning is to find a linear operator which has similar eigenvalues
to the ``true'' matrix you would like to invert (or more precisely, to $LU$-factorize), but which can be inverted faster.
If a good preconditioner can be found, the number of GMRES iterations is greatly reduced.  Many schemes for preconditioning exist,
but the version adopted in \sfincs~is to explicity form and $LU$-factorize a preconditioning matrix which is similar
to the true matrix (but somewhat simpler).  There is a basic trade-off:
the more similar the preconditioner matrix is to the true matrix, the fewer iterations will be required, but the more time will
be required to $LU$-factorize the preconditioning operator.  The usual preconditioner matrix in \sfincs~is obtained by dropping all coupling
between grid points in the speed coordinate and dropping coupling between species.  The preconditioner matrix need not be a physically
accurate or meaningful operator; as long as GMRES converges, the solution obtained will be independent of the preconditioner to whatever tolerance is specified.

\section{\sfincs~vs. \sfincsScan}
The core fortran part of \sfincs~solves the kinetic equation for each species
at a single flux surface, a single value of $E_r$, and a single set of other parameters.
However, often the goal is to determine the ambipolar $E_r$ at one or more surfaces, or to scan some other parameter.
For this task, the \sfincsScan~family of \python~scripts is available.
For a full list of the types of scans available, see section \ref{sec:sfincsScanParams}

\section{Bugs and feedback}

To report any bugs, provide feedback, or ask questions, contact Matt Landreman at
\href{mailto:matt.landreman@gmail.com}{\nolinkurl{matt.landreman@gmail.com} }

%\begin{figure}[htbp]
%\begin{center}
%%\includegraphics[width=1.0\textwidth]{poloidal_cross-sections.pdf}
%\mycaption{Trajectories of trapped particles.  (a) In tokamaks and omnigenous stellarators,
%all collisionless trajectories are confined.  (b) In contrast, trapped particles in other stellarators
%may have a radial drift that does not average to zero.
%\label{fig:poloidal_cross_sections}}
%\end{center}
%\end{figure}