### 2015-05-22 doc/manual/inputParameters.tex
Added to user manual

\chapter{Input Parameters}
\label{ch:input}

\newcommand{\param}[5]{{\setlength{\parindent}{0cm} {\ttfamily \bfseries #1}\\{\it Type}: #2\\{\it Default}: #3\\{\it When it matters}: #4\\{\it Meaning}: #5}}
\newcommand{\ssparam}[4]{{\setlength{\parindent}{0cm} {\ttfamily \bfseries #1}\\{\it Type}: #2\\{\it When it matters}: #3\\{\it Meaning}: #4}}
\newcommand{\PETScParam}[2]{{\setlength{\parindent}{0cm} {\ttfamily \bfseries #1}\\{\it Meaning}: #2}}
\newcommand{\myhrule}{{\setlength{\parindent}{0cm} \hrulefill }}

\newcommand{\true}{{\ttfamily .true.}}
\newcommand{\false}{{\ttfamily .false.}}

In this chapter we first describe all the parameters which can be included in the {\ttfamily input.namelist} file. 
Then we list some of the command-line flags associated with \PETSc~which can be useful.

\section{The {\ttfamily general} namelist}

\param{RHSMode}
{integer}
{1}
{Always}
{Option related to the number of right-hand sides (i.e. inhomogeneous drive terms) for which the kinetic equation is solved.\\

{\ttfamily RHSMode=1}: Solve for a single right-hand side.\\

{\ttfamily RHSMode=2}: Solve for 3 right-hand sides to get the 3x3 transport matrix. Presently implemented only for 1 species.\\

{\ttfamily RHSMode=3}: Solve for the 2x2 monoenergetic transport coefficients. When this option is chosen, {\ttfamily Nx} is set to 1 and only 1 species is used.
}

\myhrule

\param{outputFileName}
{string}
{``sfincsOutput.h5''}
{Always}
{Name which will be used for the HDF5 output file.  If this parameter is changed from the default value, \sfincsScan~ will not work.}

\myhrule

\param{saveMatlabOutput}
{Boolean}
{\false}
{Always}
{If this switch is set to true, Matlab m-files are created which
 store the system matrix, right-hand side, and solution vector.  If an iterative solver is used,
 the preconditioner matrix is also saved.
 PETSc usually generates an error message if you ask to save Matlab output when
the size of the linear system is more then 1400 $\times$ 1400,
 so usually this setting should be false except for very small test problems.
}

\myhrule

\param{MatlabOutputFilename}
{string}
{``sfincsMatrices''}
{Only when {\ttfamily saveMatlabOutput == .true.}}
{Start of the filenames which will be used for Matlab output.}

\myhrule

\param{saveMatricesAndVectorsInBinary}
{Boolean}
{\false}
{Always}
{If this switch is set to true, the matrix, right-hand-side, and solution of the
linear system will be saved in PETSc's binary format.  The preconditioner matrix will also
be saved if {\ttfamily tryIterativeSolver == .true.}
}

\myhrule

\param{binaryOutputFilename}
{string}
{``sfincsBinary''}
{Only when {\ttfamily saveMatricesAndVectorsInBinary == .true.}}
{Start of the filenames which will be used for binary output.}

\myhrule

\param{solveSystem}
{Boolean}
{\true}
{Always}
{If this parameter is false, the system of equations will not actually be solved.
 Sometimes it can be useful to set this parameter to {\ttfamily .false.} when debugging.
}

\section{The {\ttfamily geometryParameters} namelist}
\label{sec:geometryParameters}

\param{geometryScheme}
{integer}
{1}
{Always}
{How the magnetic geometry is specified.\\

{\ttfamily geometryScheme}==1: Use the following 3-helicity model:\\
\begin{eqnarray}
\label{eq:Bmodel}
B(\theta,\zeta)/\bar{B} 
&=& (\mbox{\ttfamily B0OverBBar}) [1 + (\mbox{\ttfamily epsilon\_t}) \cos(\theta) \\
&& + (\mbox{\ttfamily epsilon\_h}) \cos((\mbox{\ttfamily helicity\_l}) \theta - (\mbox{\ttfamily helicity\_n}) \zeta) \nonumber \\
&&  + (\mbox{\ttfamily epsilon\_antisymm}) \nonumber \\
&& \times \sin((\mbox{ \ttfamily helicity\_antisymm\_l}) \theta - (\mbox{\ttfamily helicity\_antisymm\_n}) \zeta)] \nonumber
\end{eqnarray}

{\ttfamily geometryScheme}==2: Use a 3-helicity model of the LHD standard configuration at {\ttfamily rN}=0.5.\\

{\ttfamily geometryScheme}==3: Use a 4-helicity model of the LHD inward-shifted configuration at {\ttfamily rN}=0.5.\\

{\ttfamily geometryScheme}==4: Use a 3-helicity model of the W7-X standard configuration at {\ttfamily rN}=0.5.\\

{\ttfamily geometryScheme}==5: Read the {\ttfamily vmec wout} file specified in {\ttfamily equilibriumFile} below. The file can be
       either ASCII format or \netCDF~format. (\sfincs~will auto-detect the format.).\\

{\ttfamily geometryScheme}==11: Read the IPP {\ttfamily .bc} format Boozer-coordinate file specified in {\ttfamily equilibriumFile} below. 
The file is assumed to be stellarator-symmetric.\\

{\ttfamily geometryScheme}==12: Read the IPP {\ttfamily .bc} format Boozer-coordinate file specified in {\ttfamily equilibriumFile} below. 
The file is assumed to be stellarator-asymmetric.
}

\myhrule

\param{inputRadialCoordinate}
{integer}
{3}
{When {\ttfamily geometryScheme} == 1, 5, 11, or 12}
{Which radial coordinate to use to specify the flux surface for a single calculation,
or to specify the range of flux surfaces for a radial scan. 
(When {\ttfamily geometryScheme} == 2, 3, or 4, the flux surface used will be {\ttfamily rN}=0.5.)
See section \ref{sec:radialCoordinates}
for more information about radial coordinates.\\

{\ttfamily inputRadialCoordinate}==0: Use the flux surface specified by {\ttfamily psiHat\_wish}
for a single run, and use the range specified by {\ttfamily psiHat\_min} and {\ttfamily psiHat\_max} for radial scans.\\

{\ttfamily inputRadialCoordinate}==1: Use the flux surface specified by {\ttfamily psiN\_wish}
for a single run, and use the range specified by {\ttfamily psiN\_min} and {\ttfamily psiN\_max} for radial scans.\\

{\ttfamily inputRadialCoordinate}==2: Use the flux surface specified by {\ttfamily rHat\_wish}
for a single run, and use the range specified by {\ttfamily rHat\_min} and {\ttfamily rHat\_max} for radial scans.\\

{\ttfamily inputRadialCoordinate}==3: Use the flux surface specified by {\ttfamily rN\_wish}
for a single run, and use the range specified by {\ttfamily rN\_min} and {\ttfamily rN\_max} for radial scans. \\

No matter which option you pick, the value of all 4 radial coordinates used will be saved in the output \HDF~file.
}

\myhrule

\param{inputRadialCoordinateForGradients}
{integer}
{2}
{Whenever {\ttfamily RHSMode}==1.}
{Which radial coordinate is used to use to specify the input gradients of density, temperature, and electrostatic potential,
i.e. which radial coordinate is used in the denominator of these derivatives.
See section \ref{sec:radialCoordinates}
for more information about radial coordinates.\\

{\ttfamily inputRadialCoordinateForGradients}==0: Density gradients are specified by {\ttfamily dnHatdpsiHats},
temperature gradients are specified by {\ttfamily dTHatdpsiHats}, a single $E_r$ is specified by {\ttfamily dPhiHatdpsiHat},
and the range of an $E_r$ scan is specified by {\ttfamily dPhiHatdpsiHatMin-dPhiHatdpsiHatMax}.\\

{\ttfamily inputRadialCoordinateForGradients}==1: Density gradients are specified by {\ttfamily dnHatdpsiNs},
temperature gradients are specified by {\ttfamily dTHatdpsiNs}, a single $E_r$ is specified by {\ttfamily dPhiHatdpsiN},
and the range of an $E_r$ scan is specified by {\ttfamily dPhiHatdpsiNMin-dPhiHatdpsiNMax}.\\

{\ttfamily inputRadialCoordinateForGradients}==2: Density gradients are specified by {\ttfamily dnHatdrHats},
temperature gradients are specified by {\ttfamily dTHatdrHats}, a single $E_r$ is specified by {\ttfamily dPhiHatdrHat},
and the range of an $E_r$ scan is specified by {\ttfamily dPhiHatdrHatMin-dPhiHatdrHatMax}.\\

{\ttfamily inputRadialCoordinateForGradients}==3: Density gradients are specified by {\ttfamily dnHatdrNs},
temperature gradients are specified by {\ttfamily dTHatdrNs}, a single $E_r$ is specified by {\ttfamily dPhiHatdrN},
and the range of an $E_r$ scan is specified by {\ttfamily dPhiHatdrNMin-dPhiHatdrNMax}.\\

No matter which option you pick, the gradients with respect to all 4 radial coordinates will be saved in the output \HDF~file.
}

\myhrule

\param{psiHat\_wish}
{real}
{-1}
{Only when {\ttfamily inputRadialCoordinate} == 0 and {\ttfamily geometryScheme} == 1, 5, 11, or 12.}
{Requested flux surface for the computation.  See section \ref{sec:radialCoordinates}
for more information about radial coordinates.}

\myhrule

\param{psiN\_wish}
{real}
{0.25}
{Only when {\ttfamily inputRadialCoordinate} == 1 and {\ttfamily geometryScheme} == 1, 5, 11, or 12.}
{Requested flux surface for the computation.  See section \ref{sec:radialCoordinates}
for more information about radial coordinates.}

\myhrule

\param{rHat\_wish}
{real}
{-1}
{Only when {\ttfamily inputRadialCoordinate} == 2 and {\ttfamily geometryScheme} == 1, 5, 11, or 12.}
{Requested flux surface for the computation.  See section \ref{sec:radialCoordinates}
for more information about radial coordinates.}

\myhrule

\param{rN\_wish}
{real}
{0.5}
{Only when {\ttfamily inputRadialCoordinate} == 3 and {\ttfamily geometryScheme} == 1, 5, 11, or 12.}
{Requested flux surface for the computation.  See section \ref{sec:radialCoordinates}
for more information about radial coordinates.}

\myhrule

\param{B0OverBBar}
{real}
{1.0}
{Only when {\ttfamily geometryScheme} == 1. Otherwise, {\ttfamily B0OverBBar} will be set according to the requested {\ttfamily geometryScheme}.}
{Magnitude of the (0,0) Boozer harmonic of the magnetic field strength, normalized by $\bar{B}$.}

\myhrule

\param{GHat}
{real}
{??}
{Only when {\ttfamily geometryScheme} == 1. Otherwise, {\ttfamily GHat} will be set according to the requested {\ttfamily geometryScheme}.}
{$G$ is $(c/2)\times$ the poloidal current outside the flux
 surface. Equivalently, $G$ is the coefficient of $\nabla\zeta_B$ in the
 covariant representation of $\vect{B}$ in terms of Boozer coordinates $(\theta_B,\zeta_B)$:
\begin{equation}
\label{eq:covariant}
\vect{B}(\psi,\theta_B,\zeta_B) = \beta(\psi,\theta_B,\zeta_B)\nabla\psi + I(\psi)\nabla\theta_B + G(\psi)\nabla\zeta_B.
\end{equation}
{\ttfamily GHat} is $G$ normalized by $\bar{B}\bar{R}$.}

\myhrule

\param{IHat}
{real}
{0.0}
{Only when {\ttfamily geometryScheme} == 1. Otherwise, {\ttfamily IHat} will be set according to the requested {\ttfamily geometryScheme}.}
{$I$ is $(c/2)\times$ the toroidal current inside the flux
 surface. Equivalently, $I$ is the coefficient of $\nabla\theta_B$ in the
 covariant representation of $\vect{B}$ in terms of Boozer coordinates $(\theta_B,\zeta_B)$ in (\ref{eq:covariant}).
{\ttfamily IHat} is $I$ normalized by $\bar{B}\bar{R}$.}

\myhrule

\param{iota}
{real}
{0.0}
{Only when {\ttfamily geometryScheme} == 1. Otherwise, {\ttfamily iota} will be set according to the requested {\ttfamily geometryScheme}.}
{Rotational transform, equivalent to $1/q$ where $q$ is the safety factor.}

\myhrule

\param{epsilon\_t}
{real}
{??}
{Only when {\ttfamily geometryScheme} == 1.}
{Toroidal variation in $B$, as defined by (\ref{eq:Bmodel}).}

\myhrule

\param{epsilon\_h}
{real}
{??}
{Only when {\ttfamily geometryScheme} == 1.}
{Helical variation in $B$, as defined by (\ref{eq:Bmodel}).}

\myhrule

\param{epsilon\_antisymm}
{real}
{0.0}
{Only when {\ttfamily geometryScheme} == 1.}
{Stellarator-antisymmetric variation in $B$, as defined by (\ref{eq:Bmodel}).}

\myhrule

\param{helicity\_l}
{integer}
{??}
{Only when {\ttfamily geometryScheme} == 1.}
{Poloidal mode number of the helical variation in $B$, as defined by (\ref{eq:Bmodel}).}

\myhrule

\param{helicity\_n}
{integer}
{??}
{Only when {\ttfamily geometryScheme} == 1.}
{Toroidal mode number of the helical variation in $B$, as defined by (\ref{eq:Bmodel}).}

\myhrule

\param{helicity\_antisymm\_l}
{integer}
{??}
{Only when {\ttfamily geometryScheme} == 1.}
{Poloidal mode number of the stellarator-antisymmetric variation in $B$, as defined by (\ref{eq:Bmodel}).}

\myhrule

\param{helicity\_antisymm\_n}
{integer}
{??}
{Only when {\ttfamily geometryScheme} == 1.}
{Toroidal mode number of the stellarator-antisymmetric variation in $B$, as defined by (\ref{eq:Bmodel}).
Note that you can create an up-down asymmetric tokamak by setting {\ttfamily helicity\_antisymm\_n}=0, {\ttfamily epsilon\_h}=0,
and {\ttfamily epsilon\_antisymm}$>$0.}

\myhrule

\param{psiAHat}
{real}
{??}
{Only when {\ttfamily geometryScheme} == 1. Otherwise, {\ttfamily psiAHat} will be set according to the requested {\ttfamily geometryScheme}.}
{{\ttfamily psiAHat} = $\psi_a / (\bar{B} \bar{R}^2)$
where $2 \pi \psi_a$ is the toroidal flux at the last closed flux surface.}

\myhrule

\param{aHat}
{real}
{??}
{Only when {\ttfamily geometryScheme} == 1. Otherwise, {\ttfamily aHat} will be set according to the requested {\ttfamily geometryScheme}.}
{The effective minor radius at the last closed flux surface, in units of $\bar{R}$.
The code only uses {\ttfamily aBar} for converting between the various radial coordinates in input and output quantities.
}

\myhrule

\param{equilibriumFile}
{string}
{``''}
{Only when {\ttfamily geometryScheme} == 5, 11, or 12.}
{Filename from which to load the magnetic equilibrium, either in {\ttfamily vmec wout} ASCII or \netCDF~format, or IPP {\ttfamily .bc} format.}

\myhrule

\param{VMECRadialOption}
{integer}
{1}
{Only when {\ttfamily geometryScheme} == 5.}
{Controls whether the nearest available flux surface in the {\ttfamily vmec wout} file is used, or whether radial interpolation is applied to the \vmec~data
to obtain the magnetic field components on the exact surface requested.\\

{\ttfamily VMECRadialOption}=0: Use the exact {\ttfamily XXX\_wish} flux surface requested, by interpolating from the \vmec~radial grid.\\

{\ttfamily VMECRadialOption}=1: Use a surface that may be slightly different from {\ttfamily XXX\_wish} to get the nearest available flux surface from \vmec's HALF grid.  
   The components of $\vect{B}$ in \vmec~are stored on the half grid, so interpolation is then unnecessary.\\

{\ttfamily VMECRadialOption}=2: Use a surface that may be slightly different from {\ttfamily XXX\_wish} to get the nearest available flux surface from \vmec's FULL grid.
I'm not sure why you would want this, but the feature is implemented for completeness.
}

\myhrule

\param{min\_Bmn\_to\_load}
{real}
{0.0}
{Only when {\ttfamily geometryScheme} == 5, 11, or 12.}
{Filters the magnetic field read from an input file.  Only Fourier modes $(m,n)$ for which $B_{m,n}$ is at least {\ttfamily min\_Bmn\_to\_load} will be included.}

\section{The {\ttfamily speciesParameters}~namelist}

\param{Zs}
{1D array of reals}
{1.0}
{Always}
{Charges of each species, in units of the proton charge $e$}

\myhrule

\param{mHats}
{1D array of reals}
{1.0}
{Always}
{Masses of each species, in units of the reference mass $\bar{m}$}

\myhrule

\param{nHats}
{1D array of reals}
{1.0}
{Whenever {\ttfamily RHSMode} == 1}
{Densities of each species, in units of the reference density $\bar{n}$}

\myhrule

\param{THats}
{1D array of reals}
{1.0}
{Whenever {\ttfamily RHSMode} == 1}
{Temperatures of each species, in units of the reference temperature $\bar{T}$}

\myhrule

\param{dnHatdpsiHats}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 0}
{Radial density gradients of each species, with respect to the radial coordinate $\hat{\psi}$, normalized by the reference density $\bar{n}$.}

\myhrule

\param{dTHatdpsiHats}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 0}
{Radial temperature gradients of each species, with respect to the radial coordinate $\hat{\psi}$, normalized by the reference temperature $\bar{T}$.}

\myhrule

\param{dnHatdpsiNs}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 1}
{Radial density gradients of each species, with respect to the radial coordinate $\psi_N$, normalized by the reference density $\bar{n}$.}

\myhrule

\param{dTHatdpsiNs}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 1}
{Radial temperature gradients of each species, with respect to the radial coordinate $\psi_N$, normalized by the reference temperature $\bar{T}$.}

\myhrule

\param{dnHatdrHats}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 2}
{Radial density gradients of each species, with respect to the radial coordinate $\hat{r}$, normalized by the reference density $\bar{n}$.}

\myhrule

\param{dTHatdrHats}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 2}
{Radial temperature gradients of each species, with respect to the radial coordinate $\hat{r}$, normalized by the reference temperature $\bar{T}$.}

\myhrule

\param{dnHatdrNs}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 3}
{Radial density gradients of each species, with respect to the radial coordinate $r_N$, normalized by the reference density $\bar{n}$.}

\myhrule

\param{dTHatrNs}
{1D array of reals}
{0.0}
{Whenever {\ttfamily RHSMode} == 1 and {\ttfamily inputRadialCoordinateForGradients} == 3}
{Radial temperature gradients of each species, with respect to the radial coordinate $r_N$, normalized by the reference temperature $\bar{T}$.}

\section{The {\ttfamily physicsParameters}~namelist}

\section{Directives for \sfincsScan}
\label{sec:sfincsScanParams}

The parameters for \sfincsScan~ begin with the code {\ttfamily !ss} and so are not read by the fortran part of \sfincs.
These parameters matter only when \sfincsScan~is called and are all ignored when \sfincs~is executed directly.
These parameters can appear anywhere in the {\ttfamily input.namelist} file, in any namelist or outside of any namelist.
Note that \sfincsScan~ parameters do not have defaults, unlike fortran namelist parameters.\\

\ssparam{scanType}
{integer}
{Any time \sfincsScan~is called.}
{Which type of scan will be run when \sfincsScan~is called. \\

{\ttfamily scanType=1}:  Convergence scan. (Scan the parameters in the resolutionParameters namelist.)\\

{\ttfamily scanType=2}:  Scan of $E_r$.\\

{\ttfamily scanType=3}:  Scan any one input parameter that takes a numeric value.\\

{\ttfamily scanType=4}:  Scan radius, taking the density and temperature profiles from the {\ttfamily profiles} file.
In this type of scan, the same radial electric field is used at every radius.
See {\ttfamily utils/profiles.XXX} for examples.\\

{\ttfamily scanType=5}:  Scan radius, and at each radius, scan $E_r$.  Density and temperature profiles are again
 taken from the {\ttfamily profiles} file; see {\ttfamily utils/profiles.XXX} for examples.\\

{\ttfamily scanType=21}: Read in a list of requested runs from a file {\ttfamily runspec.dat}. See {\ttfamily utils/sfincsScan\_21} for an example file.
}

\section{\PETSc~commands}

Command-line flags can be used to modify the behavior of any \PETSc~application, including \sfincs.
There are hundreds of \PETSc~options, and a list can be obtained by running with the command-line flag
{\ttfamily -help}. Here we list some of the more useful options.\\

\PETScParam{-help}
{Dumps a list of available command-line options to stdout.}

\myhrule

\PETScParam{-ksp\_view}
{Dumps detailed information to stdout related to the linear solver.}

\myhrule

\PETScParam{-ksp\_gmres\_restart {\normalfont \ttfamily$<$integer$>$}}
{After how many iterations will GMRES restart. Default is 2000. The convergence of GMRES slows every time a restart occurs, but restarts also free up memory.}

\myhrule

\PETScParam{-mat\_mumps\_icntl\_4 {\normalfont \ttfamily$<$integer$>$}}
{How much diagnostic information will be printed by \mumps. Default is 0. Set to 2 or 3 to print out useful diagnostic information about the memory required for factorizing the preconditioner.}