### 2015-05-22 doc/manual/runs.tex

\chapter{Specifying and running a computation}

\section{Normalizations}

{\setlength{\parindent}{0cm}
Dimensional quantities in \sfincs~are normalized to ``reference'' values that are denoted by a bar:\\
$\bar{B}$ = reference magnetic field, typically 1 Tesla.\\
$\bar{R}$ = reference length, typically 1 meter.\\
$\bar{n}$ = reference density, typically $10^{19}$ m$^{-3}$, $10^{20}$ m$^{-3}$, or something similar.\\
$\bar{m}$ = reference mass, typically either the mass of hydrogen or deuterium.\\
$\bar{T}$ = reference temperature in energy units, typically 1 eV or 1 keV.\\
$\bar{v} = \sqrt{2 \bar{T} / \bar{m}}$ = thermal speed at the reference temperature and mass\\
$\bar{\Phi}$ = reference electrostatic potential, typically 1 V or 1 kV.\\
}

You can choose any reference parameters you like, not just the values
suggested here. However, if you use a {\ttfamily vmec} or {\ttfamily .bc} magnetic equilibrium
by choosing {\ttfamily geometryScheme} = 5, 11, or 12, then you MUST use $\bar{B}$ = 1 Tesla and $\bar{R}$ = 1 meter.
The code ``knows'' about the reference values only through
the 3 combinations {\ttfamily Delta}, {\ttfamily alpha}, and {\ttfamily nu\_n}
in the {\ttfamily physicsParameters} namelist.

Normalized quantities are denoted by a ``hat''.  Taking the magnetic field as an example,
$\hat{B}=B/\bar{B}$, where $\hat{B}$ is called {\ttfamily BHat} in the fortran code and \HDF~output file.

\section{Radial coordinates}
\label{sec:radialCoordinates}

A variety of flux-surface label coordinates are used in other codes and in the literature.
One common choice (used in \vmec) is $\psi_N$, the toroidal flux normalized to its
value at the last closed flux surface.  Another common choice is an ``effective normalized minor radius''
$r_N$, defined by $r_N=\sqrt{\psi_N}$.  For gradients of density, temperature, and electrostatic potential (i.e. the radial
electric field), it is useful to use a dimensional local minor radius $r = r_N a$, where $a$ is some
measure of the plasma effective outer minor radius.  Finally, one could also use $\psi$ directly.
For maximum flexibility, \sfincs~permits any of these four radial coordinates to be used, and different radial
coordinates can be used in different aspects of a given computation.  Output quantities which depend
on the radial coordinate, such as radial fluxes, are often given with respect to all radial coordinates.
In \sfincs, the four radial coordinates are named as follows:\\

{\setlength{\parindent}{0cm}

{\ttfamily psiHat} = $\hat\psi$ is the toroidal flux (divided by $2\pi$), normalized by $\bar{B}\bar{R}^2$.\\

{\ttfamily psiN} = $\psi_N$ is the toroidal flux normalized by its value at the last closed flux surface.\\

{\ttfamily rHat} = $\hat{r}$ is defined as {\ttfamily aHat}$\sqrt{\mbox{\ttfamily psiN}}$, where {\ttfamily aHat} is an effective minor radius of the last closed flux surface normalized by $\bar{R}$.\\

{\ttfamily rN} = $r_N$ is defined as $\sqrt{\mbox{\ttfamily psiN}}$.\\

}

These four radial coordinates are identified by the numbers 0, 1, 2, and 3 respectively.

When setting up a run, you can make several independent choices for radial coordinates.  One parameter
you select is {\ttfamily inputRadialCoordinateForGradients} in the {\ttfamily geometryParameters} namelist.
This parameter controls which coordinate is used to specify the gradients. The possible values of {\ttfamily inputRadialCoordinateForGradients} are:\\

{\setlength{\parindent}{0cm}

0: Use derivatives with respect to {\ttfamily psiHat}: Density gradients are specified using {\ttfamily dnHatdpsiHats}, 
temperature gradients are specified using {\ttfamily dTHatdpsiHats},  a single $E_r$ is specified using
{\ttfamily dPhiHatdpsiHat}, and a range of $E_r$ for a scan is specified using {\ttfamily dPhiHatdpsiHatMin}-{\ttfamily dPhiHatdpsiHatMax}.
\\

1: Use derivatives with respect to {\ttfamily psiN}: Density gradients are specified using {\ttfamily dnHatdpsiNs}, 
temperature gradients are specified using {\ttfamily dTHatdpsiNs},  a single $E_r$ is specified using
{\ttfamily dPhiHatdpsiN}, and a range of $E_r$ for a scan is specified using {\ttfamily dPhiHatdpsiNMin}-{\ttfamily dPhiHatdpsiNMax}.
\\

2: Use derivatives with respect to {\ttfamily rHat}: Density gradients are specified using {\ttfamily dnHatdrHats}, 
temperature gradients are specified using {\ttfamily dTHatdrHats},  a single $E_r$ is specified using
{\ttfamily dPhiHatdrHat}, and a range of $E_r$ for a scan is specified using {\ttfamily dPhiHatdrHatMin}-{\ttfamily dPhiHatdrHatMax}.
\\

3: Use derivatives with respect to {\ttfamily rN}: Density gradients are specified using {\ttfamily dnHatdrNs}, 
temperature gradients are specified using {\ttfamily dTHatdrNs},  a single $E_r$ is specified using
{\ttfamily dPhiHatdrN}, and a range of $E_r$ for a scan is specified using {\ttfamily dPhiHatdrNMin}-{\ttfamily dPhiHatdrNMax}.
\\

}

Another choice involving radial coordinates is how to specify the flux surface for the computation.
This choice is made using the parameter {\ttfamily inputRadialCoordinate} in the {\ttfamily geometryParameters}
namelist, which is again an integer from 0 to 3, and this parameter need not be the same as {\ttfamily inputRadialCoordinateForGradients}.
An extra complication with specifying the flux surface is that the magnetic equilibrium file will contain data on a finite number of surfaces,
and you may wish to use one of these surfaces.  For this reason, the parameters for specifying the flux surface have {\ttfamily \_wish}
appended to the name. In other words, the allowed values for {\ttfamily inputRadialCoordinate} are:\\

{\setlength{\parindent}{0cm}

0: Specify the flux surface using {\ttfamily psiHat\_wish}.\\

1: Specify the flux surface using {\ttfamily psiN\_wish}.\\

2: Specify the flux surface using {\ttfamily rHat\_wish}.\\

3: Specify the flux surface using {\ttfamily rN\_wish}.\\

}

When using {\ttfamily geometryScheme} == 11 or 12, \sfincs~will always shift the ``wish'' value so it matches an available surface in the magnetic equilibrium file.
For {\ttfamily geometryScheme} == 5, the {\ttfamily VMECRadialOption} parameter lets you can choose whether to shift to the nearest surface in the magnetic equilibrium file,
or to interpolate the \vmec~data onto the exact value of radius you specify.

If you perform a radial scan, then there is a third choice you can make: which radial coordinate to use in the {\ttfamily profiles} file.
This choice is made with an integer 0, 1, 2, or 3 in the first non-comment line of the {\ttfamily profiles} file.
The radial coordinate used in the {\ttfamily profiles} file need not be the same as either
{\ttfamily inputRadialCoordinate} or {\ttfamily inputRadialCoordinateForGradients}.
Note however that the maximum and minimum radial electric field specified in the {\ttfamily profiles}
file must be defined as the derivative of the electrostatic potential with respect to the radial coordinate {\ttfamily inputRadialCoordinateForGradients}.

For more details about the behavior of {\ttfamily inputRadialCoordinate}, {\ttfamily inputRadialCoordinateForGradients}, and {\ttfamily VMECRadialOption},
see section \ref{sec:geometryParameters}.

\section{Trajectory models}

As discussed in \cite{sfincsPaper},
one of the capabilities of \sfincs~is to compare various models for the terms in the kinetic equation involving $E_r$.
These variations of the kinetic equation are called ``trajectory models'' in \cite{sfincsPaper}.
The relevant terms in the kinetic equation can be turned off and on by certain Boolean parameters in the {\ttfamily physicsParameters} namelist.
The models described in \cite{sfincsPaper} are selected as follows:\\

{\setlength{\parindent}{0cm}

\underline{Full trajectories:}\\
{\ttfamily 
includeXDotTerm = .true.\\
includeElectricFieldTermInXiDot = .true.\\
useDKESExBDrift = .false.\\
}

\underline{Partial trajectories:}\\
{\ttfamily
includeXDotTerm = .false.\\
includeElectricFieldTermInXiDot = .false.\\
useDKESExBDrift = .false.\\
}

\underline{DKES trajectories:}\\
{\ttfamily
includeXDotTerm = .false.\\
includeElectricFieldTermInXiDot = .false.\\
useDKESExBDrift = .true.\\
}
}

\section{Parallelization}
Choosing the number of nodes and procs.

\section{Issues with running on 1 processor}