### 2015-06-04 doc/manual/resolution.tex

\chapter{Numerical resolution parameters}
\label{ch:resolution}

Results from \sfincs~should only be believed if you are confident they
are converged with respect to the numerical resolution parameters
\Ntheta, \Nzeta, \Nxi, and \Nx.  That is, you want to be sure the
physics output of the code does not change significantly when any of
these parameters are increased. The values of \Ntheta, \Nzeta, \Nxi,
and \Nx~required for convergence depend strongly on the magnetic
geometry and collisionality. It is strongly recommended that you test for
convergence with respect to \Ntheta, \Nzeta, \Nxi, and \Nx~whenever
beginning \sfincs~calculations for a new scenario.

Note that ``convergence'' in this sense (convergence with respect to resolution parameters assuming the discretized
system is solved exactly) is separate from the convergence of GMRES/KSP.

\section{Relatively unimportant resolution parameters}

There are several resolution parameters which are almost never the limiting factor
for convergence, and so which almost never need to be adjusted.
These parameters and good values for them are \NL = 4, 
\parlink{solverTolerance} = $10^{-6}$,
\parlink{xMax} = 5.0,
and \parlink{NxPotentialsPerVth} = 40.0.  
(These values are set as the defaults.)
The latter two of these parameters are in fact ignored
for the recommended and default \parlink{xGridScheme} setting, 5.

\section{General suggestions}

The time and memory requirements of the code increase significantly
when \Ntheta, \Nzeta, \Nxi, or \Nx~are increased. Therefore, you
probably want to only scan one of these four parameters at a time
(rather than increasing two or more of them simultaneously) when
testing for convergence. (This recommended approach is the one taken
in \sfincsScan~automated convergence scans, discussed in section
\ref{sec:convergence})

When the mean-free-path is shorter than the parallel length scale of the equilibrium,
%({\ttfamily nuPrime} $\ge$ 1), 
the parameters required for convergence
do not depend much on collisionality. In the opposite
limit in which the mean-free-path is longer than the parallel length scale of the equilibrium,
%({\ttfamily nuPrime} $\le$ 1)
values of
\Nzeta~and \Nxi~required for convergence increase dramatically
as collisionality decreases. The required value of \Ntheta~increases as well, but often
not quite as dramatically. The \Nx~required for convergence does not depend
nearly as much on collisionality.  The reason is that
at low collisionality, a boundary layer develops in the distribution function
along the boundary between trapped and passing (untrapped) particles.
The location of this boundary depends on $\theta$, $\zeta$, and $\xi$,
and so high resolution is required in these coordinates to resolve the boundary layer.
But the location of the trapped-passing boundary is independent
of $x$, and hence the resolution requried in $x$ is not particularly high.

Typically you can expect to use \Nx = 5-8, with 5 being sufficient at low collisionality
and 8 being required at high collisionality.
The \Nx~required for convergence may need to increase slightly with the number of species.

The resolution parameters do not need to vary much with the radial
electric field as long as the electric field is below about 1/3 of the resonant value.
(In the notation of \cite{sfincsPaper}, when $E_*<1/3$).
For almost all experimentally relevant situations (except for HSX where $T_i/T_e$ is extremely small),
the electric field is far below the resonance, in which case you should not need to vary the resolution
parameters with the electric field.
However, if you do approach the $E_r$ resonance, \Nx~will likely need to be increased.

\section{Convergence testing}
\label{sec:convergence}

\begin{figure}
\begin{center}
\includegraphics[width=1.0\textwidth]{sfincsConvergenceScan.pdf}
\mycaption{Plot generated by \sfincsScanPlot showing a resolution convergence scan for the example
{\ttfamily geometryScheme4\_2species\_noEr}.
\label{fig:convergence}}
\end{center}
\end{figure}

To check how well the results of \sfincs~are converged with respect to the resolution
parameters, run \sfincsScan~with \parlink{scanType} = 1.
In this type of scan, a ``base case'' is first run at the
values of \Ntheta, \Nzeta, \Nxi, and \Nx~specified in the input file.
Then, each of these parameters will be varied in turn, holding the other parameters fixed.
The range of \Ntheta~in the scan is specified by the parameters \parlink{NthetaMinFactor},
\parlink{NthetaMaxFactor}, and \parlink{NthetaNumRuns}. 
Each of these three parameters is read by \sfincsScan~rather than by \sfincs~itself,
and so it must be prefaced by {\ttfamily !ss} in the input namelist file.
The ranges of \Nzeta, \Nxi, and \Nx~are set by parameters with analogous names,
as detailed in section \ref{sec:scanType1Parameters}.

For example, suppose you initially run \sfincsScan~with \Ntheta = 15, 
\parlink{NthetaMinFactor} = 0.7, \parlink{NthetaMaxFactor} = 1.5, and 
\parlink{NthetaNumRuns} = 5.
This will generate \sfincs~runs at \Ntheta = 11, 13, 15 (the base case), 19, and 23.
Notice that \sfincsScan~automatically ensures that only odd values are used.
(Only odd values of \Ntheta~and \Nzeta~are used internally in \sfincs.)
The maximum and minimum values for the scan, 11 and 23, are the nearest odd integers
to \Ntheta$\times$\parlink{NthetaMinFactor} and \Ntheta$\times$\parlink{NthetaMaxFactor} respectively.
Notice also that there are `gaps' at  \Ntheta = 17 and 21 since \sfincsScan~attempts to space the values
logarithmically rather than uniformly.

In addition to the 4 resolution parameters above, \sfincsScan~also allows the parameters
\parlink{solverTolerance}, \parlink{xMax}, and \parlink{NxPotentialsPerVth}
to be scanned.  These latter two parameters are not used for the default \parlink{xGridScheme}
so it is usually unnecessary to scan them.  It is also usually unnecessary to scan \parlink{solverTolerance}
since the default value is usually robust.

A directory will be created for each run.  Each directory will contain
a copy of {\ttfamily input.namelist} in which \Ntheta, \Nzeta, \Nx, or \Nxi~has been
altered appropriately by \sfincsScan. Each directory will also contain a job file for the run.

You do not need to scan all variables. For example, if you do not wish to scan \Nx,
you can either set \parlink{NxNumRuns} = 0, or you can not specify
\parlink{NxNumRuns} in the {\ttfamily input.namelist} file.

Once the individual runs in the scan have begun to finish, you can plot the results by running \sfincsScanPlot.
You can plot the results before all the runs in the scan finish, although at least the base case
run must be completed.  Different quantities will be displayed depending on
\parlink{RHSMode} and \parlink{includePhi1}.

Typical convergence behavior is illustrated in figure \ref{fig:convergence},
showing the results of \sfincsScan~and \sfincsScanPlot~for the example
{\ttfamily geometryScheme4\_2species\_noEr}.
In this figure, a very large number of values are included for the resolution
parameters, more than you would likely include in routine use of the code.
The red vertical lines emphasize the ``base case'' resolution parameters.
For all parameters scanned, the physical output quantities (particle flux, heat flux, etc)
are essentially unchanged when any of the resolution parameters
are increased beyond the base case value.  Thus, the results in the base case are converged,
and so we can believe the results.
Observe in the figure that the output quantities tend to oscillate when \Nxi~is incremented by 1 and insufficient
\Nxi~is used. Results for even and odd \Nxi~converge to the same value, as one would hope.

If you wish to add more runs to the scan at a later time,
you can alter the relevant \sfincsScan~parameters in the original {\ttfamily input.namelist}~file
and run \sfincsScan~again from the original directory.  The \sfincsScan~code will automatically detect which runs
have already been submitted, so duplicate runs will not be generated.
For example, consider the scan of \Ntheta~described previously in this section,
and suppose you wish to fill in the gaps at \Ntheta = 17 and 21 in the original scan.
To do this, you could set \parlink{NthetaNumRuns} to a large number like 100
and re-run \sfincsScan.  Since \sfincsScan~intelligently avoids duplication,
the only new runs generated will be  \Ntheta = 17 and 21.

If you are ultimately intending to scan $E_r$, you probably only need to run a convergence scan
at a single value of $E_r$.  This is because the resolution requirements of \sfincs~are not sensitive to
$E_r$, as discussed above.
Also, if you are ultimately intending to run \sfincs~at a range of minor radii,
it is reasonably to only run a convergence scan at a single radius close to the magnetic axis.
This is because the typically peaked shape of the temperature profile means that
collisionality is lowest near the axis.  Since the resolution requirements of \sfincs~are most
demanding at low collisionality, then if the code is converged at the radius of lowest collisionality,
it should be converged at all radii.

\section{Typical resolution requirements}