### 2016-07-20 doc/manual/version3/equations.tex
Beginning version 4

\chapter{Kinetic equations}

The \sfincs~code is capable of solving many different variants of the drift-kinetic equation.
In this section we summarize the most common drift-kinetic equations
which can be solved with the code, giving the associated input parameters
(which are all in the  {\ttfamily \hyperref[sec:physicsParameters]{physicsParameters}} namelist.)
For other choices of these input parameters, 
other combinations of terms in the drift-kinetic equation can be used than those
given here.
For more information, see section \ref{sec:trajectoryModels}, section \ref{sec:qn}, section \ref{sec:magneticDrifts},
and the version 3 technical documentation.
The following terms are always included in \sfincs:
parallel streaming, the magnetic mirror force, the collision operator,
and the inhomogeneous drive term from radial gradients.
Other terms generally can be turned on and off as desired using parameters in the
{\ttfamily \hyperref[sec:physicsParameters]{physicsParameters}} namelist.

Throughout this section, gradients and other partial derivatives are taken
at constant $x_a = v/\sqrt{2T_a/m_a}$ and at constant $\xi = v_{||} / v$, unless denoted otherwise with subscripts.
We use $s$ to denote the species, $r$ to denote any radial coordinate, (expressions are independent
of the specific radial coordinate used,)
$C_a$ to denote the collision operator, and $S_a$ to denote the source-sink term.
A more detailed discussion of the equations implemented in the code can be found
in the technical documentation for \sfincs~version 3, available in the {\ttfamily /doc} directory
of the repository.

\section{Default: Full $\vect{E}\times\vect{B}$ trajectories; no poloidal or toroidal magnetic drifts; flux function potential}

\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r \right) \cdot \nabla f_{s1} \\
&&+ \left[ - \frac{(1-\xi^2)v}{2B} \nabla_{||} B
+\frac{(1-\xi^2)\xi}{2B^3} \frac{d\Phi_0}{dr}\vect{B}\times\nabla r\cdot\nabla B \right]
 \frac{\partial f_{s1}}{\partial \xi} \nonumber \\
&&-(\vect{v}_{ma} \cdot\nabla r) \frac{Z_s e}{2 T_s x_s} \frac{d\Phi_0}{dr} \frac{\partial f_{a1}}{\partial x_s} \nonumber \\
&&+ (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
 = C_a + S_a \nonumber
\end{eqnarray}
Note that this equation is equivalent to the following one, in which the independent variables
are $(\mu,v_{||})$ instead of $(\xi,x_s)$:
\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r \right) \cdot (\nabla f_{s1})_{\mu, v_{||}} \\
&&+ \left[ - \mu \nabla_{||} B
-\frac{v_{||}}{B^2} \frac{d\Phi_0}{dr} \vect{b}\times\nabla B \times \nabla r \right]
\left( \frac{\partial f_{a1}}{\partial v_{||}} \right)_{\mu} \nonumber \\
&& + (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
=C_a + S_a \nonumber
\end{eqnarray}
These equivalent forms of the kinetic equation are selected using \\
\parlink{includeXDotTerm} = \true  \;\;\; (Default) \\
\parlink{includeElectricFieldTermInXiDot} = \true \;\;\; (Default) \\
\parlink{useDKESExBDrift} = \false \;\;\; (Default) \\
\parlink{magneticDriftScheme} = 0 \;\;\; (Default) \\
\parlink{includePhi1} = \false \;\;\; (Default) \\
\parlink{includeRadialExBDrive} = \false \;\;\; (Default) \\
\parlink{nonlinear} = \false \;\;\; (Default)

\section{DKES $\vect{E}\times\vect{B}$ trajectories; no poloidal or toroidal magnetic drifts; flux function potential}

This form of the kinetic equation is useful for benchmarking with DKES and other codes
that use the same equation.

\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{\left< B^2\right>} \vect{B}\times\nabla r \right) \cdot \nabla f_{s1} \\
&& - \frac{(1-\xi^2)v}{2B} (\nabla_{||} B)
 \frac{\partial f_{s1}}{\partial \xi} \nonumber \\
&&+ (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
 = C_a + S_a \nonumber
\end{eqnarray}
This form of the kinetic equation is selected using \\
\parlink{includeXDotTerm} = \false  \;\;\; (Not default) \\
\parlink{includeElectricFieldTermInXiDot} = \false \;\;\; (Not default) \\
\parlink{useDKESExBDrift} = \true \;\;\; (Not default) \\
\parlink{magneticDriftScheme} = 0 \;\;\; (Default) \\
\parlink{includePhi1} = \false \;\;\; (Default) \\
\parlink{includeRadialExBDrive} = \false \;\;\; (Default) \\
\parlink{nonlinear} = \false \;\;\; (Default)

\section{Full $\vect{E}\times\vect{B}$ trajectories; including poloidal and toroidal magnetic drifts; flux function potential}

\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r  + \vect{v}_{ms}\right) \cdot \nabla \theta \frac{\partial f_{s1}}{\partial \theta} \\
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r  + \vect{v}_{ms}\right) \cdot \nabla \zeta \frac{\partial f_{s1}}{\partial \zeta} \nonumber \\
&&+ \left\{ - \frac{(1-\xi^2)v}{2B} \nabla_{||} B
+\frac{(1-\xi^2)\xi}{2B^3} \frac{d\Phi_0}{dr}\vect{B}\times\nabla r\cdot\nabla B \right. \nonumber \\
&& \left. \hspace{0.25in} -\frac{T_s x_s^2 (\nabla r\cdot\nabla\theta\times\nabla\zeta)}{Z_s e B^3} (1-\xi^2)\xi
\left[ \left( \frac{\partial B_r}{\partial \zeta} - \frac{\partial B_\zeta}{\partial r}\right)\frac{\partial B}{\partial\theta}
+ \left(\frac{\partial B_\theta}{\partial r} - \frac{\partial B_r}{\partial\theta}\right) \frac{\partial B}{\partial\zeta}\right]
\right\}
 \frac{\partial f_{s1}}{\partial \xi} \nonumber \\
&&-(\vect{v}_{ma} \cdot\nabla r) \frac{Z_s e}{2 T_s x_s} \frac{d\Phi_0}{dr} \frac{\partial f_{a1}}{\partial x_s} \nonumber \\
&&+ (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
 = C_a + S_a \nonumber
\end{eqnarray}
Notice the magnetic drifts affect the coefficients of $\partial f_{s1}/\partial \theta$, $\partial f_{s1}/\partial \zeta$, 
and $\partial f_{s1}/\partial \xi$,
but there is no change to the coefficient of $\partial f_{s1}/\partial x_s$.
This form of the kinetic equation is selected using \\
\parlink{includeXDotTerm} = \true  \;\;\; (Default) \\
\parlink{includeElectricFieldTermInXiDot} = \true \;\;\; (Default) \\
\parlink{useDKESExBDrift} = \false \;\;\; (Default) \\
\parlink{magneticDriftScheme} = 1 \;\;\; (Not default) \\
\parlink{includePhi1} = \false \;\;\; (Default) \\
\parlink{includeRadialExBDrive} = \false \;\;\; (Default) \\
\parlink{nonlinear} = \false \;\;\; (Default)

\section{Full $\vect{E}\times\vect{B}$ trajectories; no poloidal or toroidal magnetic drifts; leading $\Phi_1$ term}

In this form of the drift-kinetic equation, we include the largest term involving $\Phi_1$,
associated with the adiabatic response.
However, other terms considered in Ref \cite{Regana2013} are not included.

\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r \right) \cdot \nabla f_{s1} \\
&&+ \left[ - \frac{(1-\xi^2)v}{2B} \nabla_{||} B
+\frac{(1-\xi^2)\xi}{2B^3} \frac{d\Phi_0}{dr}\vect{B}\times\nabla r\cdot\nabla B \right]
 \frac{\partial f_{s1}}{\partial \xi} \nonumber \\
&&-(\vect{v}_{ma} \cdot\nabla r) \frac{Z_s e}{2 T_s x_s} \frac{d\Phi_0}{dr} \frac{\partial f_{a1}}{\partial x_s} \nonumber \\
&&+\frac{Z_s e}{T_s} f_{sM} v_{||} \nabla_{||} \Phi_1 \nonumber \\
&&+ (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
 = C_a + S_a \nonumber
\end{eqnarray}
Note that this equation is equivalent to the following one, in which the independent variables
are $(\mu,v_{||})$ instead of $(\xi,x_s)$:
\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r \right) \cdot (\nabla f_{s1})_{\mu, v_{||}} \\
&&+ \left[ - \mu \nabla_{||} B
-\frac{v_{||}}{B^2} \frac{d\Phi_0}{dr} \vect{b}\times\nabla B \times \nabla r \right]
\left( \frac{\partial f_{a1}}{\partial v_{||}} \right)_{\mu} \nonumber \\
&&+\frac{Z_s e}{T_s} f_{sM} v_{||} \nabla_{||} \Phi_1 \nonumber \\
&& + (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
=C_a + S_a \nonumber
\end{eqnarray}
These equivalent forms of the kinetic equation are selected using \\
\parlink{includeXDotTerm} = \true  \;\;\; (Default) \\
\parlink{includeElectricFieldTermInXiDot} = \true \;\;\; (Default) \\
\parlink{useDKESExBDrift} = \false \;\;\; (Default) \\
\parlink{magneticDriftScheme} = 0 \;\;\; (Default) \\
\parlink{includePhi1} = \true \;\;\; (Not default) \\
\parlink{includeRadialExBDrive} = \false \;\;\; (Default) \\
\parlink{nonlinear} = \false \;\;\; (Default)

\section{Full $\vect{E}\times\vect{B}$ trajectories; no poloidal or toroidal magnetic drifts; Garc\'{i}a-Rega\~{n}a $\Phi_1$ terms}

This form of the drift-kinetic equation is nearly identical to equation (11) in Ref \cite{Regana2013}.
The one difference (which should be small) is that at the end of (11), Garc\'{i}a-Rega\~{n}a
has a term $\propto \vect{v}_m \cdot \nabla \Phi_1$, which is not (yet) in \sfincs.

\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r \right) \cdot \nabla f_{s1} \\
&&+ \left[ - \frac{(1-\xi^2)v}{2B} \nabla_{||} B
+\frac{(1-\xi^2)\xi}{2B^3} \frac{d\Phi_0}{dr}\vect{B}\times\nabla r\cdot\nabla B 
-\frac{Z_s e}{v m_s}(1-\xi^2)(\nabla_{||}\Phi_1)
\right]
 \frac{\partial f_{s1}}{\partial \xi} \nonumber \\
&&+ \left[
-(\vect{v}_{ma} \cdot\nabla r) \frac{Z_s e}{2 T_s x_s} \frac{d\Phi_0}{dr} 
-\frac{Z_s e \xi}{v_s m_s}
\right] \frac{\partial f_{a1}}{\partial x_s} \nonumber \\
&&+\frac{Z_s e}{T_s} f_{sM} v_{||} \nabla_{||} \Phi_1 \nonumber \\
&& + (\vect{v}_E \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr}  + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM} \nonumber \\
&&+ (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
 = C_a + S_a \nonumber
\end{eqnarray}
Note that this equation is equivalent to the following one, in which the independent variables
are $(\mu,v_{||})$ instead of $(\xi,x_s)$:
\begin{eqnarray}
&&\left(v_{||}\vect{b} + \frac{d\Phi_0}{dr} \frac{1}{B^2} \vect{B}\times\nabla r \right) \cdot (\nabla f_{s1})_{\mu, v_{||}} \\
&&+ \left[ -\frac{Z_s e}{m_s} \nabla_{||} \Phi_1
- \mu \nabla_{||} B
-\frac{v_{||}}{B^2} \frac{d\Phi_0}{dr} \vect{b}\times\nabla B \times \nabla r \right]
\left( \frac{\partial f_{a1}}{\partial v_{||}} \right)_{\mu} \nonumber \\
&&+\frac{Z_s e}{T_s} f_{sM} v_{||} \nabla_{||} \Phi_1 \nonumber \\
&& + (\vect{v}_E \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr}  + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM} \nonumber \\
&& + (\vect{v}_{ma} \cdot \nabla r) \left[ \frac{1}{n_a} \frac{dn_a}{dr} + \frac{Z_s e}{T_s} \frac{d\Phi_0}{dr} + \left(x_s^2-\frac{3}{2}\right) \frac{1}{T_s} \frac{dT_s}{dr}\right] f_{aM}
=C_a + S_a \nonumber
\end{eqnarray}
These equivalent forms of the kinetic equation are selected using \\
\parlink{includeXDotTerm} = \true  \;\;\; (Default) \\
\parlink{includeElectricFieldTermInXiDot} = \true \;\;\; (Default) \\
\parlink{useDKESExBDrift} = \false \;\;\; (Default) \\
\parlink{magneticDriftScheme} = 0 \;\;\; (Default) \\
\parlink{includePhi1} = \true \;\;\; (Not default) \\
\parlink{includeRadialExBDrive} = \true \;\;\; (Not default) \\
\parlink{nonlinear} = \true \;\;\; (Not default)