### 2018-02-09 doc/manual/version3/outputParameters.tex
Working on updating the manual

\chapter{Output Parameters}
\label{ch:output}

In this chapter we describe output parameters which are available in the \sfincs~output file \parlink{outputFileName} after a successful \sfincs~run. 
The output is in the \HDF~format, and the output can be looked at e.g. by using the {\ttfamily h5dump} command on Linux. 
Note that all parameters in the output file are case-sensitive.

\section{Convergence flags}

\section{Output fluxes and flows}

%The default values are usually best for the parameters in this namelist.
%
%\myhrule
%
%\param{outputFileName}
%{string}
%{``sfincsOutput.h5''}
%{Always}
%{Name which will be used for the HDF5 output file.  If this parameter is changed from the default value, \sfincsScan~ will not work.}
%
%\myhrule