### 2018-02-15 doc/manual/version3/outputParameters.tex
Working on manual

\chapter{Output Parameters}
\label{ch:output}

\newcommand{\outputparam}[4]{{\setlength{\parindent}{0cm} {\ttfamily \bfseries \hypertarget{#1}{#1}}\\{\it Type}: #2\\{\it Available when}: #3\\{\it Meaning}: #4}}
\newcommand{\outputparamWithUnderscore}[5]{{\setlength{\parindent}{0cm} {\ttfamily \bfseries \hypertarget{#2}{#1}}\\{\it Type}: #3\\{\it Available when}: #4\\{\it Meaning}: #5}}
\newcommand{\underscoreparlink}[2]{{\ttfamily \hyperlink{#2}{#1}}}

In this chapter we describe output parameters which are available in the \sfincs~output file \parlink{outputFileName} after a %successful 
\sfincs~run. 
The output is in the \HDF~format, and the output can be looked at e.g. by using the {\ttfamily h5dump} command on Linux. 
Note that all parameters in the output file are case-sensitive. 
Moreover, many of the input parameters in chapter~\ref{ch:input} also exist in the output file and we refrain from repeating them here.

\section{Convergence and test parameters}

\myhrule

\outputparam{finished}
{integer}
{\sfincs~has finished.}
{If this variable exists, then \sfincs~reached the end of all requested computations and exited gracefully. 
Note that this does not guarantee that a nonlinear calculation (\parlink{includePhi1} == \true) has converged.}

\myhrule

\outputparam{integerToRepresentTrue}
{integer}
{Always}
{Since \HDF~does not have a Boolean datatype, this integer value is used to represent \true.}

\myhrule

\outputparam{integerToRepresentFalse}
{integer}
{Always}
{Since \HDF~does not have a Boolean datatype, this integer value is used to represent \false.}

\myhrule

\outputparam{didNonlinearCalculationConverge}
{integer}
{\parlink{RHSMode} == 1, \parlink{solveSystem} == \true, \parlink{includePhi1} == \true and \parlink{finished} available.}
{If this variable is \true~the nonlinear iteration has converged. Interpret this Boolean data by comparing to \parlink{integerToRepresentTrue}/\parlink{integerToRepresentFalse}.}

\myhrule

\outputparam{lambda}
{real}
{\parlink{RHSMode} == 1, \parlink{solveSystem} == \true and \parlink{includePhi1} == \true.}
{Lagrange multiplier associated with the constraint that $\left\langle \Phi_1\right\rangle = 0$, which is included in the quasineutrality equation. %Should be within machine precision of 0.
This parameter should come out as negligible in quasineutrality, e.g. its size should be significantly smaller than the maximum charge density of the plasma species $Z_s n_s$.
\todo{Is this correct?}}

\myhrule

\outputparam{NIterations}
{integer}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{Number of iterations in the nonlinear calculation. (In a linear calculation this will be 1.)}

\myhrule

\section{Radial fluxes and flows}

A detailed description of the most important output fluxes and flows are given in the Technical Documentation for \sfincs. 
Here we give a concise version of some of the quantities. 
Transforming the dimension of the output quantities into SI units is typically done with the normalizations 
$\bar{B}$, $\bar{R}$, $\bar{n}$, $\bar{m}$, $\bar{T}$, $\bar{v}$, $\bar{\Phi}$ (see section~\ref{sec:normalizations}).

\subsubsection{Radial fluxes}

The radial fluxes have the form 
\[
\left\langle \bm{X} \cdot \nabla Y \right\rangle
\]
where $Y$ is a flux-surface label and $\left< \ldots \right>$ denotes a flux surface average. In \sfincs~there are 4 possible labels: $\hat{\psi}$, $\psi_N$, $\hat{r}$ and $r_N$ (see section~\ref{sec:radialCoordinates}), indicated by the suffix {\ttfamily \_psiHat}, {\ttfamily \_psiN}, {\ttfamily \_rHat} or {\ttfamily \_rN} in the output variable name. 
Below we will only present the description of the {\ttfamily XXX\_psiHat} variables, but the description of the other variables are readily obtained by noting that 
all flux-surface labels are dimensionless which implies that the normalization factor will be the same. E.g. when $\displaystyle \left\langle \bm{X} \cdot \nabla \hat{\psi} \right\rangle = \text{\ttfamily NormalizationFactor} \cdot \text{\ttfamily XXX\_psiHat}$ then also $\displaystyle \left\langle \bm{X} \cdot \nabla \hat{r} \right\rangle = \text{\ttfamily NormalizationFactor} \cdot \text{\ttfamily XXX\_rHat}$ applies with the same {\ttfamily NormalizationFactor}.

For the radial flux of each quantity (particle, momentum and energy), and for each radial coordinate, there is an output quantity associated with the zeroth-order $f_{s0}$ or with the full distribution function $f_s = f_{s0} + f_{s1}$, and there is also a quantity associated with the radial magnetic drift, with the radial $\bm{E} \times \bm{B}$-drift or with the total drift $\bm{v}_{ds} = \bm{v}_{ms} + \bm{v}_{E}$. Here we only present all possible combinations for the particle flux, but the pattern is the same for the momentum and energy flux.

\myhrule

\outputparamWithUnderscore{particleFlux\_vd\_psiHat}
{particleFlux_vd_psiHat}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1, \parlink{solveSystem} == \true and \parlink{includePhi1} == \true.}
{$\displaystyle \text{\ttfamily particleFlux\_vd\_psiHat} = \frac{\bar{R}}{\bar{n} \bar{v}} \left\langle \int d^3v f_s \, \bm{v}_{ds} \cdot \nabla \hat{\psi}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}

\myhrule

\outputparamWithUnderscore{particleFlux\_vm\_psiHat}
{particleFlux_vm_psiHat}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily particleFlux\_vm\_psiHat} = \frac{\bar{R}}{\bar{n} \bar{v}} \left\langle \int d^3v f_s \, \bm{v}_{ms} \cdot \nabla \hat{\psi}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}

\myhrule

\outputparamWithUnderscore{particleFlux\_vE\_psiHat}
{particleFlux_vE_psiHat}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1, \parlink{solveSystem} == \true and \parlink{includePhi1} == \true.}
{$\displaystyle \text{\ttfamily particleFlux\_vE\_psiHat} = \frac{\bar{R}}{\bar{n} \bar{v}} \left\langle \int d^3v f_s \, \bm{v}_{E} \cdot \nabla \hat{\psi}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}

\myhrule

\outputparamWithUnderscore{particleFlux\_vm0\_psiHat}
{particleFlux_vm0_psiHat}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily particleFlux\_vm0\_psiHat} = \frac{\bar{R}}{\bar{n} \bar{v}} \left\langle \int d^3v f_{s0} \, \bm{v}_{ms} \cdot \nabla \hat{\psi}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration. This variable should cancel \underscoreparlink{particleFlux\_vE0\_psiHat}{particleFlux_vE0_psiHat} in the last nonlinear iteration since $f_{s0}$ cannot drive radial transport (i.e. in a linear calculations this quantity should be 0 to high precision unless there is a radial current in the magnetic equilibrium).}

\myhrule

\outputparamWithUnderscore{particleFlux\_vE0\_psiHat}
{particleFlux_vE0_psiHat}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1, \parlink{solveSystem} == \true and \parlink{includePhi1} == \true.}
{$\displaystyle \text{\ttfamily particleFlux\_vE0\_psiHat} = \frac{\bar{R}}{\bar{n} \bar{v}} \left\langle \int d^3v f_{s0} \, \bm{v}_{E} \cdot \nabla \hat{\psi}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration. This variable should cancel \underscoreparlink{particleFlux\_vm0\_psiHat}{particleFlux_vm0_psiHat} in the last nonlinear iteration since $f_{s0}$ cannot drive radial transport.}

\myhrule

\outputparamWithUnderscore{particleFlux\_vd1\_psiHat}
{particleFlux_vd1_psiHat}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1, \parlink{solveSystem} == \true and \parlink{includePhi1} == \true.}
{$\displaystyle \text{\ttfamily particleFlux\_vd1\_psiHat} = \text{\ttfamily particleFlux\_vm\_psiHat}~+~\text{\ttfamily particleFlux\_vE0\_psiHat}$. The rationale for defining this quantity is that (at least for $E_r = 0$) it gives the particle flux that would be expected if \parlink{includePhi1} == \false.}

\myhrule

\outputparamWithUnderscore{particleFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}}
{particleFlux_vXN_Y}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{See above for various $\textcolor{magenta}{XN}$.}
{For different options of $\textcolor{magenta}{XN}$ see output {\ttfamily particleFlux} parameters defined above. 
For different options of $\textcolor{magenta}{Y}$ see flux-surface labels (section~\ref{sec:radialCoordinates}). 
$\displaystyle \text{\ttfamily particleFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}} = \frac{\bar{R}}{\bar{n} \bar{v}} \left\langle \int d^3v f_{s\textcolor{magenta}{N}} \, \bm{v}_{\textcolor{magenta}{X}s} \cdot \nabla \textcolor{magenta}{Y}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}

\myhrule

%\outputparamWithUnderscore{momentumFlux\_vm\_psiHat}
%{momentumFluxvmpsiHat}
%{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
%{Always}
%{$\displaystyle \text{\ttfamily momentumFlux\_vm\_psiHat} = \frac{\bar{R}}{\bar{n} \bar{v}^2 \bar{B} \bar{m}} \left\langle \int d^3v f_s \, B m_s v_{\|} \bm{v}_{ms} \cdot \nabla \hat{\psi}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}
%
%\myhrule

\outputparamWithUnderscore{momentumFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}}
{momentumFlux_vXN_Y}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{See above for various $\textcolor{magenta}{XN}$.}
{For different options of $\textcolor{magenta}{XN}$ see  
\underscoreparlink{particleFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}}{particleFlux_vXN_Y}.
For different options of $\textcolor{magenta}{Y}$ see flux-surface labels (section~\ref{sec:radialCoordinates}). 
$\displaystyle \text{\ttfamily momentumFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}} = \frac{\bar{R}}{\bar{n} \bar{v}^2 \bar{B} \bar{m}} \left\langle \int d^3v f_{s\textcolor{magenta}{N}} \, B m_s v_{\|} \bm{v}_{\textcolor{magenta}{X}s} \cdot \nabla \textcolor{magenta}{Y}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}

\myhrule

\outputparamWithUnderscore{heatFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}}
{heatFlux_vXN_Y}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{See above for various $\textcolor{magenta}{XN}$.}
{For different options of $\textcolor{magenta}{XN}$ see  
\underscoreparlink{particleFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}}{particleFlux_vXN_Y}.
For different options of $\textcolor{magenta}{Y}$ see flux-surface labels (section~\ref{sec:radialCoordinates}). 
$\displaystyle \text{\ttfamily heatFlux\_v\textcolor{magenta}{XN}\_\textcolor{magenta}{Y}} = \frac{\bar{R}}{\bar{n} \bar{v}^3 \bar{m}} \left\langle \int d^3v f_{s\textcolor{magenta}{N}} \, \frac{m_s v^2}{2} \bm{v}_{\textcolor{magenta}{X}s} \cdot \nabla \textcolor{magenta}{Y}\right\rangle$ for all plasma species. The final result is in the last index corresponding to the last nonlinear iteration.}

\myhrule

\outputparam{flow}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily flow} =  \frac{1}{\bar{n} \bar{v}} \int d^3v \, v_{\|} f_s$.}

\myhrule

\outputparam{velocityUsingFSADensity}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily velocityUsingFSADensity} =  \parlink{flow}~/~\parlink{nHats}.}

\myhrule

\outputparam{velocityUsingTotalDensity}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily velocityUsingTotalDensity} =  \parlink{flow}~/~\parlink{totalDensity}.}

\myhrule

\outputparam{MachUsingFSAThermalSpeed}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily MachUsingFSAThermalSpeed} =  (\parlink{mHats}~/~\parlink{THats})${}^{1/2}$~$\cdot$~\parlink{flow}~/~\parlink{nHats}.}

\myhrule

\outputparam{FSABFlow}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily FSABFlow} =  \frac{1}{\bar{v} \bar{B} \bar{n}} \left\langle B \int d^3v \, v_{\|} f_s \right\rangle$.}

\myhrule

\outputparam{FSABVelocityUsingFSADensity}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily FSABVelocityUsingFSADensity} =  \parlink{FSABFlow}~/~\parlink{nHats}.}

\myhrule

\outputparam{FSABVelocityUsingFSADensityOverB0}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily FSABVelocityUsingFSADensityOverB0} =  ($\bar{B}$~/~$B_0$)~$\cdot$~\parlink{FSABFlow}~/~\parlink{nHats}.}

\myhrule

\outputparam{FSABVelocityUsingFSADensityOverRootFSAB2}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily FSABVelocityUsingFSADensityOverRootFSAB2} =  (1~/~\parlink{nHats})~$\cdot$~\parlink{FSABFlow}~/~(\parlink{FSABHat2})${}^{1/2}$.}

\myhrule

\outputparam{jHat}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily jHat} = \frac{\bm{j} \cdot \bm{b}}{e \bar{n} \bar{v}} = \frac{1}{e \bar{n} \bar{v}} \sum_s Z_s e \int d^3v \, v_{\|} f_s = \sum_s Z_s \cdot \text{\parlink{flow}}_s$.}

\myhrule

\outputparam{FSABjHat}
{$\text{\ttfamily NIterations}$ array of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily FSABjHat} =  \sum_s Z_s \cdot \text{\parlink{FSABFlow}}_s$.}

\myhrule

\outputparam{FSABjHatOverB0}
{$\text{\ttfamily NIterations}$ array of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily FSABjHatOverB0} =  \frac{\bar{B}}{B_0} \text{\parlink{FSABjHat}}$.}

\myhrule

\outputparam{FSABjHatOverRootFSAB2}
{$\text{\ttfamily NIterations}$ array of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily FSABjHatOverRootFSAB2} =  \frac{\bar{\text{\parlink{FSABjHat}}}}{\left(\text{\parlink{FSABHat2}}\right)^{1/2}}$.}

\myhrule

\section{Distributions}

\myhrule

\outputparam{densityPerturbation}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily densityPerturbation} =  \frac{1}{\bar{n}} \int d^3v f_{s1}$.}

\myhrule

\outputparam{totalDensity}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily totalDensity} = \parlink{nHats}~+~\parlink{densityPerturbation}.}

\myhrule

\outputparam{FSADensityPerturbation}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily FSADensityPerturbation} =  \frac{1}{\bar{n}} \left\langle \int d^3v f_{s1} \right\rangle$. 
Should be within machine precision of 0.}

\myhrule

\outputparam{pressurePerturbation}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily densityPerturbation} =  \frac{1}{\bar{n} \bar{T}} \frac{m_s}{3} \int d^3v \, v^2 f_{s1}$.}

\myhrule

\outputparam{totalPressure}
{$\text{\ttfamily Nzeta} \times \text{\ttfamily Ntheta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{{\ttfamily totalPressure} = \parlink{nHats} $\cdot$ \parlink{THats}~+~\parlink{pressurePerturbation}.}

\myhrule

\outputparam{FSAPressurePerturbation}
{$\text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{RHSMode} == 1 and \parlink{solveSystem} == \true.}
{$\displaystyle \text{\ttfamily FSADensityPerturbation} =  \frac{1}{\bar{n} \bar{T}} \frac{m_s}{3} \left\langle \int d^3v \, v^2 f_{s1} \right\rangle$. 
Should be within machine precision of 0.}

\myhrule

\outputparam{delta\_f}
{$M_x \times M_{\xi} \times M_{\zeta} \times M_{\theta} \times \text{\ttfamily Nspecies} \times \text{\ttfamily NIterations}$ matrix of reals}
{\parlink{export\_delta\_f} == \true.}
{?}

\myhrule

\section{Miscellaneous}

\myhrule

\outputparam{VPrimeHat}
{real}
{Always}
{$\displaystyle \text{\ttfamily VPrimeHat} =  \int_{0}^{2 \pi} d \theta \int_{0}^{2 \pi} d \zeta \frac{\bar{B}}{\bar{R}} \frac{1}{\nabla \psi \cdot \nabla \theta \times \nabla \zeta}$.}

\myhrule

\outputparam{FSABHat2}
{real}
{Always}
{$\displaystyle \text{\ttfamily FSABHat2} =  \frac{1}{\bar{B}^2} \left\langle B^2 \right\rangle$.}

\myhrule

%The default values are usually best for the parameters in this namelist.
%
%\myhrule
%
%\param{outputFileName}
%{string}
%{``sfincsOutput.h5''}
%{Always}
%{Name which will be used for the HDF5 output file.  If this parameter is changed from the default value, \sfincsScan~ will not work.}
%
%\myhrule