### 2019-01-17 doc/classical_radial_fluxes_2019-01-17.tex
Adds a radial heat flux calculation to the classicalTransport module. Unifies the classical heat flux and particle flux by writing both in terms of Braginskii matrices. Adds documentation of the classical flux calculation.

\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath}
%\usepackage{relsize}

%\usepackage{fixmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[usenames,dvipsnames]{color}
%\usepackage[small,font=it]{caption}
\usepackage{amssymb}
%\usepackage{icomma}
%\usepackage{hyperref}
%\usepackage{mcode}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{units}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\footnotesize \parbox{11cm}{Braginskii matrix approach to classical transport from the linearized Fokker-Planck collision operator}}
\rhead{\footnotesize \parbox{2cm}{Stefan Buller}}
\renewcommand\headheight{24pt}
%\usepackage{dirtytalk} \say{} command for quotations

\makeatletter
\newenvironment{tablehere}
  {\def\@captype{table}}
  {}

\newenvironment{figurehere}
  {\def\@captype{figure}}
  {}

\newsavebox\myboxA
\newsavebox\myboxB
\newlength\mylenA

\newcommand*\obar[2][0.75]{% OverBAR, adds bar over an element
    \sbox{\myboxA}{$\m@th#2$}%
    \setbox\myboxB\null% Phantom box
    \ht\myboxB=\ht\myboxA%
    \dp\myboxB=\dp\myboxA%
    \wd\myboxB=#1\wd\myboxA% Scale phantom
    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
    \addtolength\mylenA{-\the\wd\myboxB}%
    \ifdim\wd\myboxB<\wd\myboxA%
       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
    \else
        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
    \fi}

\makeatother

%\def\equationautorefname{ekvation}
%\def\tableautorefname{tabell}
%\def\figureautorefname{figur}
%\def\sectionautorefname{sektion}
%\def\subsectionautorefname{sektion}

\DeclareMathOperator\erf{erf}
%\DeclareMathOperator\mod{mod}

\newcommand{\ordo}[1]{{\cal O}\left( #1 \right)}
\DeclareMathOperator{\sgn}{sgn}
%\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\im}{\ensuremath{\mathrm{i}}}
\newcommand{\e}{\ensuremath{\mathrm{e}}}
\newcommand{\E}{\ensuremath{\mathcal{E}}}
\newcommand{\p}{\ensuremath{\partial}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}

\newcommand{\bra}[1]{\langle #1 \mid}
\newcommand{\ket}[1]{\mid #1 \rangle}
\newcommand\matris[4]{\ensuremath{\begin{pmatrix} #1 & #2 \\ #3 & #4\end{pmatrix}}}
\renewcommand{\d}{\ensuremath{\mathrm{d}}}
%\DeclareMathOperator*{\sgn}{sgn}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\lang}{\left\langle}
\newcommand{\rang}{\right\rangle}

%to get leftrightarrow over tensors of rank-2.
\def\shrinkage{2.1mu}
\def\vecsign{\mathchar"017E}
\def\dvecsign{\smash{\stackon[-1.95pt]{\mkern-\shrinkage\vecsign}{\rotatebox{180}{$\mkern-\shrinkage\vecsign$}}}}
\def\dvec#1{\def\useanchorwidth{T}\stackon[-4.2pt]{#1}{\,\dvecsign}}
\usepackage{stackengine}
\stackMath
\def\perfect{\textsc{Perfect}}

\lstset{language=[90]Fortran,
  basicstyle=\ttfamily,
  keywordstyle=\color{red},
  commentstyle=\color{green},
  morecomment=[l]{!\ }% Comment only with space after !
   frame=single,
   breaklines=true,
   postbreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\color{blue}\hookrightarrow\space}}
}

\usepackage[
    backend=biber,
    style=alphabetic,
    citestyle=authoryear,
    natbib=true,
    url=false,
    doi=false,
    isbn=false,
    eprint=false,
    sorting=none,
    maxcitenames=2,
    maxbibnames=3
]{biblatex}

\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}
\AtEveryBibitem{\clearfield{issue}}
\AtEveryCitekey{\clearfield{issue}}

\bibliography{../../../plasma-bib}

%=================================================================
\begin{document}
\section{Drift-kinetic classical transport}
These notes derive the classical transport under the same set of assumptions as in the linearized drift-kinetic equation solved by the $\delta f$-code neoclassical code \textsc{Sfincs} (and probably most other neoclassical codes).

\subsection{Motivation}
For a mass-ratio expanded ion-impurity collision operator, we found that the classical transport of collisional impurities in W7-X can be comparable to the neoclassical transport. Thus, it would be of interest to have a general numerical tool to calculate the classical transport alongside the neoclassical transport from a drift-kinetic solver such as \textsc{Sfincs}.

\subsection{Gyrophase dependent part of $f$}
In many ways, calculating the classical transport is simpler than calculating the neoclassical transport, as the gyrophase dependent part of the distribution function is smaller than the gyrophase-independent in the expansion parameter $\rho/L$, and the gyrophase dependent part to required order is given entirely by the zeroth order gyrophase independent distribution. When $f_{a0}$ is a Maxwellian $f_{aM}$, the gyrophase dependent part is given by
\begin{equation}
\tilde{f}_{a1} = -\vec{\rho}_a(\gamma) \cdot \nabla f_{Ma} = -\vec{\rho}(\gamma) \cdot \nabla \psi  \frac{\p f_{Ma}}{\p \psi}, \label{eq:ftilde}
\end{equation}
where $\gamma$ is the gyrophase, $\psi$ a flux-label and $\vec{\rho}$ the gyrophase vector
\begin{equation}
\vec{\rho}_a = \frac{1}{\Omega_a}\vec{b} \times \vec{v}_\perp,
\end{equation}
with $\vec{v}_\perp$ the velocity perpendicular to the magnetic field, $\Omega_a=Z_a eB/m_a$ the gyrofrequency, $\vec{b}$ the unit vector in the direction of the magnetic field $\vec{B}$; and $\psi$ a flux-label, which acts as our radial coordinate.

\autoref{eq:ftilde} can be massaged a bit to yield (we drop the tilde since we will only consider the gyrophase dependent part of $f_{a1}$ here)
\begin{equation}
f_{a1} = \frac{1}{\Omega_a}(\vec{b}\times \nabla \psi) \cdot  \vec{v}_\perp\frac{\p f_{Ma}}{\p \psi},
\end{equation}
and we introduce the local coordinate system $\vec{v}_\perp = v_2 \hat{e}_2 + v_3 \hat{e}_3$, where $\hat{e}_2$, $\hat{e}_3$ form an orthonormal triplet with $\vec{b}$. There is some freedom in how we align our $\hat{e}_2$, $\hat{e}_3$ coordinates, and we choose these vectors so that $\hat{e}_2$ is entirely in the $(\vec{b}\times \nabla \psi)$ direction. Thus
\begin{equation}
f_{a1} = \frac{1}{\Omega_a}|\vec{b}\times \nabla \psi| v_2 \frac{\p f_{Ma}}{\p \psi}.
\end{equation}

Finally, we evaluate $\frac{\p f_{Ma}}{\p \psi}$. For heavy impurities, the lowest order distribution may vary on the flux-surface in response to a flux-surface variation of the electrostatic potential, in which case, we write
\begin{equation}
f_{Ma} = \eta_a(\psi) \left(\frac{m_a}{2\pi T_a}\right)^{3/2} \exp{\left(-\frac{m_av^2}{2T_a} - \frac{Z_a e \tilde{\Phi}}{T_a}\right)},
\end{equation}
where the \emph{pseudo-density} $\eta_a$ is related to the ordinary density by
\begin{equation}
\eta_a = n_a \e^{Z_a e \tilde{\Phi}/T_a}.
\end{equation}
The gradient thus becomes
\begin{equation}
\frac{\p f_{Ma}}{\p \psi} = f_{Ma}\left(\frac{\d \ln\eta_a}{\d \psi} + \frac{Z_a e}{T_a} \frac{\p \tilde{\Phi}}{\p \psi}  + \frac{Z_a e \tilde{\Phi}}{T_a} \frac{\d \ln T_a}{\d \psi}
 + \left[\frac{m_a v^2}{2T_a} - \frac{3}{2}\right]  \frac{\d \ln T_a}{\d \psi}\right). \label{eq:dpsif0}
\end{equation}
To simplify notation, we will introduce the quantities
\begin{align}
\alpha_{1a} \equiv & \frac{\d \ln\eta_a}{\d \psi} + \frac{Z_a e}{T_a} \frac{\p \tilde{\Phi}}{\p \psi}  + \frac{Z_a e \tilde{\Phi}}{T_a} \frac{\d \ln T_a}{\d \psi}
  - \frac{3}{2}  \frac{\d \ln T_a}{\d \psi} \\
  \alpha_{2a} \equiv & \frac{\d \ln T_a}{\d \psi},
\end{align}
so that $\frac{\p f_{Ma}}{\p \psi} = f_{Ma}(\alpha_{a1} + \alpha_{2a} x_a^2)$,
where
\begin{equation}
x_a^2 \equiv \frac{m_a v^2}{2T_a} \equiv \frac{v^2}{v_{Ta}^2}.
\end{equation}
Thus
\begin{equation}
f_{a1} = \frac{|\vec{b}\times \nabla \psi| }{\Omega_a}v_2 f_{Ma}(\alpha_{1a} + \alpha_{2a} x_a^2).
\end{equation}

\subsection{Radial fluxes}
The classical radial flux of particle and heat is given by
\begin{align}
  \Gamma_a \equiv \lang \vec{\Gamma}_a \cdot \nabla \psi \rang^{\text{C}} \equiv& \lang \frac{\vec{b} \times \nabla \psi}{Z_a e B} \cdot\vec{R}_a \rang,\label{eq:Gamma} \\
  Q_a \equiv \lang \vec{Q}_a \cdot \nabla \psi \rang^{\text{C}} \equiv& \lang \frac{\vec{b} \times \nabla \psi}{Z_a e B} \cdot\vec{G}_a \rang,\label{eq:Q}
\end{align}
where we have introduced the \emph{friction force} and \emph{energy-weighted friction force}
\begin{align}
  \vec{R}_a &= \int \d^3 v m_a \vec{v} C[f_{a}],\\
  \vec{G}_a &= \int \d^3 v \frac{m_a v^2}{2}  m_a \vec{v} C[f_{a}].
\end{align}
Here, $C[f_a]$ is the Fokker-Planck collision operator accounting for the effects of all species acting on $f_a$
\begin{equation}
    C[f_a] = \sum_b C_{ab}\left[f_{a},f_{b}\right].
    \end{equation}
    An important property of the Fokker-Planck operator is that it preserves the gyrophase dependence of $f_a$. Thus, only the gyrophase dependent part of $f_{a}$ will contribute to the friction force in the $\vec{b} \times \nabla\psi$ direction.
    As the gyrophase dependent part of $f_a$ acts as a small correction to $f_{Ma}$, we can linearize the Fokker-Planck operator around $f_{Ma}$ and write
\begin{equation}
  \begin{aligned}
    C[f_a] =& \sum_b C_{ab}\left[f_{Ma} + f_{1a},f_{Mb}+f_{1b}\right] \\
    \approx & \sum_b C_{ab}[f_{Ma},f_{Mb}] + C_{ab}[f_{Ma},f_{b1}] + C_{ab}[f_{a1},f_{Mb}],
      \end{aligned}
    \end{equation}
    where the gyrophase-independent $C_{ab}[f_{Ma},f_{Mb}]$ part will not contribute to the classical transport. With this result, we write the relevant components of $\vec{R}_a$ and $\vec{G}_a$ as
    \begin{align}
  \vec{b} \times \nabla \psi \cdot \vec{R}_a &= m_a  |\vec{b} \times \nabla \psi|\sum_b \int \!\d^3 v \, v_2 \left(C_{ab}[f_{a1},f_{Mb}] + C_{ab}[f_{Ma},f_{b1}]\right),\\
  \vec{b} \times \nabla \psi \cdot \vec{G}_a &= m_a T_a |\vec{b} \times \nabla \psi| \sum_b \int \!\d^3 v\, v_2 x_a^2  \left(C_{ab}[f_{a1},f_{Mb}] + C_{ab}[f_{Ma},f_{b1}] \right).
    \end{align}
    Inserting our expressions for $f_{1}$, we thus get
    \begin{align}
      &\begin{aligned}
        (\vec{b} \times \nabla \psi) \cdot \vec{R}_a = m_a  |\vec{b} \times \nabla \psi|^2\sum_b \int \!\d^3 v \, v_2 &\left( \frac{\alpha_{1a}}{\Omega_a}  C_{ab}[v_2 f_{Ma},f_{Mb}] + \frac{\alpha_{2a}}{\Omega_a} C_{ab}[v_2 x_a^2 f_{Ma} ,f_{Mb}] \right.\\
          &\left.+ \frac{\alpha_{1b}}{\Omega_b}C_{ab}[f_{Ma},v_2 f_{Mb}] + \frac{\alpha_{2b}}{\Omega_b}C_{ab}[f_{Ma},v_2 x_b^2 f_{Mb} ]\right),
      \end{aligned}\\
      &\begin{aligned}
        (\vec{b} \times \nabla \psi) \cdot \vec{G}_a = m_a T_a |\vec{b} \times \nabla \psi|^2 \sum_b \int \!\d^3 v\, v_2 x_a^2  &\left( \frac{\alpha_{1a} }{\Omega_a}C_{ab}[v_2 f_{Ma},f_{Mb}]
          + \frac{\alpha_{2a}}{\Omega_a}C_{ab}[v_2 x_a^2 f_{Ma} ,f_{Mb}]
        \right.\\
        &\left.+ \frac{\alpha_{1b}}{\Omega_b}C_{ab}[f_{Ma},v_2 f_{Ma}]
          + \frac{\alpha_{2b}}{\Omega_b}C_{ab}[f_{Ma},v_2 x_b^2f_{Ma}] 
        \right).
      \end{aligned}
    \end{align}
    These friction-force projections can be conveniently expressed in terms of Braginskii matrix elements
    \begin{align}
  M_{ab}^{jk} = \frac{\tau_{ab}}{n_a} \int v_2 L_j^{(3/2)}(x_a^2) C_{ab}\left[\frac{m_a v_2}{T_a} L_k^{(3/2)}(x_a^2) f_{a0},f_{b0}\right] \\
  N_{ab}^{jk} = \frac{\tau_{ab}}{n_a} \int v_2 L_j^{(3/2)}(x_a^2) C_{ab}\left[f_{a0},\frac{m_b v_2}{T_b} L_k^{(3/2)}(x_b^2)  f_{b0}\right],
\end{align}
where
\begin{equation}
\tau_{ab} = \frac{12 \pi^{3/2} \epsilon_0^2}{\sqrt{2} e^4 \ln \Lambda} \frac{m_a^{1/2} T_a^{3/2}}{Z_a^2 Z_b^2 n_b},
\end{equation}
and the relevant Sonine polynomials are
\begin{align}
  L_0^{(3/2)}(x_a^2) = 1 \\
  L_1^{(3/2)}(x_a^2) = \frac{5}{2} - x_a^2,
\end{align}
so that
\begin{align}
  x_a^2 - \frac{5}{2} =& -L_1^{(3/2)}(x_a^2)  \label{eq:qmotiv}\\
  x_a^2  =& \frac{5}{2}L_0^{(3/2)}(x_a^2) -L_1^{(3/2)}(x_a^2)  \\
  1 = &L_0^{(3/2)}(x_a^2).
\end{align}
\autoref{eq:qmotiv} makes it convenient to calculate the conductive heat flux $q_a = Q_a - \frac{5}{2} T_a \Gamma_a$ rather than directly calculating $Q_a$. With this in mind, we write
\begin{align}
      &\begin{aligned}
        (\vec{b} \times \nabla \psi) \cdot \vec{R}_a = m_a  |\vec{b} \times \nabla \psi|^2\sum_b \frac{n_a}{\tau_{ab}}&\left[ \frac{\alpha_{1a}}{\Omega_a}  \frac{T_a}{m_a}M_{ab}^{00} + \frac{\alpha_{2a}}{\Omega_a}\frac{T_a}{m_a} \left(\frac{5}{2}M_{ab}^{00} - M_{ab}^{01}\right)  \right.\\
          &\left.+ \frac{\alpha_{1b}}{\Omega_b}\frac{T_b}{m_b}N_{ab}^{00}+ \frac{\alpha_{2b}}{\Omega_b}\frac{T_b}{m_b}\left(\frac{5}{2} N_{ab}^{00} - N_{ab}^{01}\right)\right],
      \end{aligned}\\
      &\begin{aligned}
        (\vec{b} \times \nabla \psi) \cdot \left(\vec{G}_a - T_a \frac{5}{2}\vec{R}_a\right) = -m_a T_a |\vec{b} \times \nabla \psi|^2 \sum_b \frac{n_a}{\tau_{ab}} &\left[ \frac{\alpha_{1a} }{\Omega_a} \frac{T_a}{m_a}M_{ab}^{10}
          + \frac{\alpha_{2a}}{\Omega_a}\frac{T_a}{m_a}\left(\frac{5}{2}M_{ab}^{10} - M_{ab}^{11}\right) 
        \right.\\
        &\left.+ \frac{\alpha_{1b}}{\Omega_b}\frac{T_b}{m_b} N_{ab}^{10}
          + \frac{\alpha_{2b}}{\Omega_b}\frac{T_b}{m_b}\left(\frac{5}{2}N_{ab}^{10} - N_{ab}^{11}\right) 
        \right].
      \end{aligned}
\end{align}
The classical particle-flux and heat-flux thus become

\begin{align}
  \hspace*{-18ex}\Gamma_a =&  \frac{m_a}{Z_a  e^2}\lang |\nabla \psi|^2 \sum_b \frac{n_a n_b}{B^2 \tau_{ab} n_b}\left[ \frac{\alpha_{1a} T_a}{Z_a}  M_{ab}^{00} + \frac{\alpha_{2a} T_a}{Z_a} \left(\frac{5}{2}M_{ab}^{00} - M_{ab}^{01}\right) + \frac{\alpha_{1b} T_b}{Z_b}N_{ab}^{00}+ \frac{\alpha_{2b} T_b}{Z_b}\left(\frac{5}{2} N_{ab}^{00} - N_{ab}^{01}\right)\right] \rang \\
  \hspace*{-18ex}q_a =&  -\frac{T_a m_a}{Z_a e^2}\lang |\nabla \psi|^2 \sum_b \frac{n_a n_b}{B^2 \tau_{ab} n_b }\left[ \frac{\alpha_{1a} T_a}{Z_a}  M_{ab}^{10} + \frac{\alpha_{2a} T_a}{Z_a} \left(\frac{5}{2}M_{ab}^{10} - M_{ab}^{11}\right) + \frac{\alpha_{1b} T_b}{Z_b}N_{ab}^{10}+ \frac{\alpha_{2b} T_b}{Z_b}\left(\frac{5}{2} N_{ab}^{10} - N_{ab}^{11}\right)\right] \rang.
\end{align}

The values for the matrix-components can be found in the appendix of Helander+Sigmar. We can simplify the above expressions slightly by noting the symmetry properties
\begin{align}
  M_{ab}^{jk} &= M_{ab}^{kj}, \\
  N_{ab}^{jk} &= \frac{T_a v_{Ta}}{T_b v_{Tb}} N_{ba}^{kj}, \\
  N_{ab}^{j0} &=-M_{ab}^{j0} = -M_{ab}^{0j},  \\
  N_{ab}^{0j} &= \frac{T_a v_{Ta}}{T_b v_{Tb}} N_{ba}^{j0} =-\frac{T_a v_{Ta}}{T_b v_{Tb}} M_{ba}^{j0}  =-\frac{T_a^{3/2} m_b^{1/2}}{T_b^{3/2} m_a^{1/2}} M_{ba}^{0j}, \\
\end{align}
where the last property follows from applying the other 3. Thus, we only need
\begin{align}
  M_{ab}^{00} =& - \frac{1+\frac{m_a}{m_b}}{\left(1+\frac{m_a T_b}{m_b T_a}\right)^{3/2}} = - \frac{\left(1+\frac{m_a}{m_b}\right)\left(1+\frac{m_a T_b}{m_b T_a}\right)}{\left(1+\frac{m_a T_b}{m_b T_a}\right)^{5/2}}\\
  M_{ab}^{01} =& - \frac{3}{2} \frac{1+\frac{m_a}{m_b}}{\left(1+\frac{m_a T_b}{m_b T_a}\right)^{5/2}} \\
  M_{ba}^{01} =& - \frac{3}{2} \frac{1+\frac{m_b}{m_a}}{\left(1+\frac{m_b T_a}{m_a T_b}\right)^{5/2}} =- \frac{3}{2} \frac{\left(1+\frac{m_a}{m_b}\right)\frac{m_b}{m_a}\left(\frac{m_a T_b}{m_b T_a}\right)^{5/2}}{\left(1+\frac{m_a T_b}{m_b T_a}\right)^{5/2}} \\
  M_{ab}^{11} =& - \frac{\frac{13}{4} + 4 \frac{m_a T_b}{m_b T_a} + \frac{15}{2}\left(\frac{m_a T_b}{m_b T_a}\right)^2 }{\left(1+\frac{m_a T_b}{m_b T_a}\right)^{5/2}} \\
  N_{ab}^{11} =& \frac{27}{4}  \frac{\frac{m_a}{m_b}}{\left(1+\frac{m_a T_b}{m_b T_a}\right)^{5/2}}. \\
\end{align}

The property $N_{ab}^{j0} =-M_{ab}^{j0} = -M_{ab}^{0j}$ also means that the radial electric field does not contribute to the classical heat or particle flux.

Writing out the $\alpha_1$ and $\alpha_2$ and employing the relevant symmetry properties, we obtain the particle flux
\begin{equation}
\begin{aligned}
 \hspace*{-10ex} \Gamma_a =  \frac{m_a}{Z_a e^2}\sum_b \frac{1}{\tau_{ab} n_b}&\left[\lang  n_a n_b\frac{ |\nabla \psi|^2}{B^2} \rang M_{ab}^{00} \left(\frac{ T_a}{Z_a}  \frac{\d \ln\eta_a}{\d \psi} -\frac{ T_b}{Z_b}  \frac{\d \ln\eta_b}{\d \psi}\right) \right.\\
    &\left.+    \lang n_a n_b\frac{  |\nabla \psi|^2}{B^2} e\tilde{\Phi} \rang M_{ab}^{00} \left(\frac{\d \ln T_a}{\d \psi} -\frac{\d \ln T_b}{\d \psi}\right) \right.\\
   &\left. + \lang n_a n_b \frac{|\nabla \psi|^2 }{B^2}\rang\left(  \left(M_{ab}^{00} - M_{ab}^{01}\right)\frac{ T_a}{Z_a}\frac{\d \ln T_a}{\d \psi} 
    - \left(M_{ab}^{00} - \frac{m_a T_b}{m_b T_a} M_{ab}^{01}\right)\frac{ T_b}{Z_b}\frac{\d \ln T_b}{\d \psi}\right)\right] 
\end{aligned}
\end{equation}
and the heat-flux
\begin{equation}
\begin{aligned}
  \hspace*{-10ex} q_a =  -\frac{T_a m_a}{Z_a e^2} \sum_b \frac{1}{\tau_{ab} n_b} &\left[
    \lang n_a n_b\frac{|\nabla \psi|^2}{B^2 } \rang M_{ab}^{01}\left(\frac{T_a}{Z_a}  \frac{\d \ln\eta_a}{\d \psi} - \frac{T_b}{Z_b}\frac{\d \ln\eta_b}{\d \psi}\right) \right.\\
    &\left.+\lang n_a n_b\frac{|\nabla \psi|^2}{B^2 } e \tilde{\Phi} \rang M_{ab}^{01} \left(\frac{\d \ln T_a}{\d \psi}-\frac{\d \ln T_b}{\d \psi}\right) \right.\\
    &\left.+ \lang n_a n_b\frac{|\nabla \psi|^2}{B^2 } \rang \left(\left(M_{ab}^{01} - M_{ab}^{11}\right)\frac{ T_a}{Z_a} \frac{\d \ln T_a}{\d \psi}
    - \left(M_{ab}^{01} + N_{ab}^{11}\right)\frac{T_b}{Z_b}\frac{\d \ln T_b}{\d \psi}\right) \right]
\end{aligned}
\end{equation}

\section{SFINCS implementation}
\textsc{Sfincs} uses normalized quantities (denoted with a hat), where the normalized radial fluxes are defined in terms of the physical fluxes as
\begin{align}
  \hat{\Gamma}_a &= \frac{\bar{R}}{\bar{n}\bar{v}} \frac{1}{Z_a e} \lang B^{-2} \left(\vec{B} \times \nabla \hat{\psi}\right) \cdot \vec{R}_a \rang, \\
  \hat{Q}_a &= \frac{\bar{R}}{\bar{n}\bar{v} \bar{T}} \frac{1}{Z_a e} \lang B^{-2} \left(\vec{B} \times \nabla \hat{\psi}\right) \cdot \vec{G}_a \rang,
\end{align}
where $\hat{\psi} = \psi/(\bar{B}\hat{R}^2)$ is just a new choice of radial coordinate rather than a normalization -- see the \textsc{Sfincs} manual for more details. We also define the dimensionless parameters $\alpha = e\bar{\Phi}/\bar{T}$ and $\Delta = \frac{\sqrt{2\bar{m}\bar{T}}}{e\bar{R}\bar{B}}$, and the dimensionless collisionality $\nu_n$
\begin{align}
  \bar{\nu} &\equiv \frac{\sqrt{2}}{12 \pi^{3/2}} \frac{\bar{n} e^4 \ln \Lambda}{\epsilon_0^2 \bar{m}^{1/2} \bar{T}^{3/2}} \\
  \nu_n &\equiv \bar{\nu} \frac{\bar{R} \sqrt{\bar{m}}}{\sqrt{2\bar{T}}} =\frac{1}{12 \pi^{3/2}} \frac{\bar{n} \bar{R} e^4 \ln \Lambda}{\epsilon_0^2 \bar{T}^{2}},
\end{align}
so that
\begin{equation}
\frac{1}{\tau_{ab} n_b} = \frac{Z_a^2 Z_b^2}{\hat{m}_a^{1/2} \hat{T}_a^{3/2}} \frac{\sqrt{2} \bar{T}^{1/2}}{\bar{n} \bar{R} \bar{m}^{1/2}} \nu_n = \frac{Z_a^2 Z_b^2}{\hat{m}_a^{1/2} \hat{T}_a^{3/2}} \frac{\sqrt{2} \bar{T}^{1/2}\bar{m}^{1/2} \bar{B}}{\bar{n} \bar{R} \bar{B} \bar{m}} \nu_n = \frac{Z_a^2 Z_b^2}{\hat{m}_a^{1/2} \hat{T}_a^{3/2}} \frac{e\bar{B}}{\bar{n}\bar{m}} \Delta \nu_n.
\end{equation}
With this, the normalized particle and conductive heat-flux becomes
\begin{equation}
\begin{aligned}
  \hspace*{-10ex} \hat{\Gamma}_a =   \frac{\Delta^2 \nu_n}{2}\frac{Z_a\hat{m}_a^{1/2}}{\hat{T}_a^{3/2}}\sum_b Z_b^2 &\left[
    \lang  \hat{n}_a \hat{n}_b\frac{ |\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2} \rang M_{ab}^{00} \left(\frac{ \hat{T}_a}{Z_a}  \frac{\d \ln\eta_a}{\d \hat{\psi}} -\frac{ \hat{T}_b}{Z_b}  \frac{\d \ln\eta_b}{\d \hat{\psi}}\right) \right.\\
    &\left.+    \lang \hat{n}_a \hat{n}_b\frac{  |\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2} \hat{\Phi}_1 \rang \alpha  M_{ab}^{00} \left(\frac{\d \ln T_a}{\d \hat{\psi}} -\frac{\d \ln T_b}{\d \hat{\psi}}\right) \right.\\
   &\left. + \lang \hat{n}_a \hat{n}_b \frac{| \bar{R} \nabla \hat{\psi}|^2 }{\hat{B}^2}\rang\left(  \left(M_{ab}^{00} - M_{ab}^{01}\right)\frac{ \hat{T}_a}{Z_a}\frac{\d \ln T_a}{\d \hat{\psi}} 
    - \left(M_{ab}^{00} - \frac{\hat{m}_a \hat{T}_b}{\hat{m}_b \hat{T}_a} M_{ab}^{01}\right)\frac{ \hat{T}_b}{Z_b}\frac{\d \ln T_b}{\d \hat{\psi}}\right)\right] 
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
  \hspace*{-10ex} \hat{q}_a =  -\frac{\Delta^2 \nu_n}{2} \frac{Z_a\hat{m}_a^{1/2}}{\hat{T}_a^{1/2}} \sum_b Z_b^2 &\left[
    \lang \hat{n}_a \hat{n}_b\frac{|\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2 } \rang M_{ab}^{01}\left(\frac{\hat{T}_a}{Z_a}  \frac{\d \ln\eta_a}{\d \hat{\psi}} - \frac{\hat{T}_b}{Z_b}\frac{\d \ln\eta_b}{\d \hat{\psi}}\right) \right.\\
    &\left.+\lang \hat{n}_a \hat{n}_b\frac{|\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2 }  \hat{\Phi}_1 \rang \alpha M_{ab}^{01} \left(\frac{\d \ln T_a}{\d \hat{\psi}}-\frac{\d \ln T_b}{\d \hat{\psi}}\right) \right.\\
    &\left.+ \lang \hat{n}_a \hat{n}_b\frac{|\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2 } \rang \left(\left(M_{ab}^{01} - M_{ab}^{11}\right)\frac{ \hat{T}_a}{Z_a} \frac{\d \ln T_a}{\d \hat{\psi}}
    - \left(M_{ab}^{01} + N_{ab}^{11}\right)\frac{\hat{T}_b}{Z_b}\frac{\d \ln T_b}{\d \hat{\psi}}\right) \right].
\end{aligned}
\end{equation}
Finally, using $\hat{n}_a = \hat{\eta}_a \e^{-Z_a \alpha \hat{\Phi}_1/\hat{T}_a}$, we get
\begin{equation}
\hat{n}_a\hat{n}_b = \hat{\eta}_a \hat{\eta}_b \exp{\left(-\alpha\hat{\Phi}_1\left[\frac{Z_a}{\hat{T}_a} + \frac{Z_b}{\hat{T}_b}\right]\right)},
\end{equation}
where $\hat{\eta}$ is the ``density'' in \textsc{Sfincs}, which is a flux-function. In the code, we define
\begin{align}
  G_{ab}^{(1)} =& \lang \exp{\left(-\alpha\hat{\Phi}_1\left[\frac{Z_a}{\hat{T}_a} + \frac{Z_b}{\hat{T}_b}\right]\right)}\frac{|\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2 } \rang \\
  G_{ab}^{(2)} =& \lang \exp{\left(-\alpha\hat{\Phi}_1\left[\frac{Z_a}{\hat{T}_a} + \frac{Z_b}{\hat{T}_b}\right]\right)}\frac{|\bar{R}\nabla \hat{\psi}|^2}{\hat{B}^2 } \hat{\Phi}_1 \rang,
\end{align}
so that
\begin{equation}
\begin{aligned}
  \hspace*{-10ex} \hat{\Gamma}_a =   \frac{\Delta^2 \nu_n}{2}\frac{Z_a\hat{m}_a^{1/2} \hat{\eta}_a}{\hat{T}_a^{3/2}}\sum_b Z_b^2 \hat{\eta}_b&\left[
    G_{ab}^{(1)} M_{ab}^{00} \left(\frac{ \hat{T}_a}{Z_a}  \frac{\d \ln\eta_a}{\d \hat{\psi}} -\frac{ \hat{T}_b}{Z_b}  \frac{\d \ln\eta_b}{\d \hat{\psi}}\right) \right.\\
    &\left.+    G_{ab}^{(2)} \alpha  M_{ab}^{00} \left(\frac{\d \ln T_a}{\d \hat{\psi}} -\frac{\d \ln T_b}{\d \hat{\psi}}\right) \right.\\
   &\left. +G_{ab}^{(1)}\left(  \left(M_{ab}^{00} - M_{ab}^{01}\right)\frac{ \hat{T}_a}{Z_a}\frac{\d \ln T_a}{\d \hat{\psi}} 
    - \left(M_{ab}^{00} - \frac{\hat{m}_a \hat{T}_b}{\hat{m}_b \hat{T}_a} M_{ab}^{01}\right)\frac{ \hat{T}_b}{Z_b}\frac{\d \ln T_b}{\d \hat{\psi}}\right)\right],
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
  \hspace*{-10ex} \hat{q}_a =  -\frac{\Delta^2 \nu_n}{2} \frac{Z_a\hat{m}_a^{1/2} \hat{\eta}_a}{\hat{T}_a^{1/2}} \sum_b Z_b^2 \hat{\eta}_b &\left[
    G_{ab}^{(1)}  M_{ab}^{01}\left(\frac{\hat{T}_a}{Z_a}  \frac{\d \ln\eta_a}{\d \hat{\psi}} - \frac{\hat{T}_b}{Z_b}\frac{\d \ln\eta_b}{\d \hat{\psi}}\right) \right.\\
    &\left.+G_{ab}^{(2)} \alpha M_{ab}^{01} \left(\frac{\d \ln T_a}{\d \hat{\psi}}-\frac{\d \ln T_b}{\d \hat{\psi}}\right) \right.\\
    &\left.+ G_{ab}^{(1)} \left(\left(M_{ab}^{01} - M_{ab}^{11}\right)\frac{ \hat{T}_a}{Z_a} \frac{\d \ln T_a}{\d \hat{\psi}}
    - \left(M_{ab}^{01} + N_{ab}^{11}\right)\frac{\hat{T}_b}{Z_b}\frac{\d \ln T_b}{\d \hat{\psi}}\right) \right],
\end{aligned}
\end{equation}
and
\begin{equation}
\hat{Q}_a = \hat{q}_a + \frac{5}{2} \hat{T}_a \hat{\Gamma}_a. 
\end{equation}

\paragraph{Thanks to Sarah Newton} for bringing to my attention that the classical transport can be calculated using the Braginskii matrices.

\end{document}